---
title: "WGCNA GO Analysis"
author: "Kristen Brown, modified ZD"
date: "1/29/2024"
output: html_document
---

# WGCNA GO Analysis

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
sessionInfo() #provides list of loaded packages and version of R. 
```

```{r}
#BiocManager::install("simplifyEnrichment")
```

First, load the necessary packages.

```{r load_libraries, include = TRUE}
library(goseq)
library(dplyr)
library(forcats)
library(ggplot2)
library(gridExtra)
library(tidyr)
library(grDevices)
library(reshape2)
library(Rmisc)
library(ggpubr)
library(tibble)
library(gridExtra)
library(tidyr)
library(zoo)
library(ComplexHeatmap)
library(circlize)
library(GSEABase)
library(data.table)
library(stringr)
library(GenomicRanges)
library(rtracklayer)
library(rrvgo)
library(simplifyEnrichment)
library(scales)
```

## Load in data
```{r}
library(rtracklayer)
gff<-rtracklayer::import("../../data/Pocillopora_acuta_HIv2.genes_fixed.gff3")
gff<-as.data.frame(gff) %>% dplyr::select(-Parent)
dim(gff) # 478988     9
names(gff) 

transcripts <- subset(gff, type == "transcript")
transcripts_gr <- makeGRangesFromDataFrame(transcripts, keep.extra.columns=TRUE) #extract length information
transcript_lengths <- width(transcripts_gr) #isolate length of each gene
seqnames<-transcripts_gr$ID #extract list of gene id 
lengths<-cbind(seqnames, transcript_lengths)
lengths<-as.data.frame(lengths) #convert to data frame

dim(transcripts) #33730    13
```

```{r}
kegg <- read.delim("../../data/Pocillopora_acuta_HIv2.genes.KEGG_results.txt",header = FALSE)
kegg <- as.data.frame(kegg)
colnames(kegg)[1] <- "gene_id" 
colnames(kegg)[2] <- "KEGG_new"
head(kegg)
```


```{r}
eggnog<-read.delim("../../data/Pocillopora_acuta_HIv2.genes.EggNog_results.txt")#this file contains all of the go terms, descriptions, kegg, etc
eggnog<- plyr::rename(eggnog, c("X.query"="gene_id"))
head(eggnog,2)
```

```{r}
gogene <- merge(transcripts, eggnog, by=c("gene_id"), all=T)
gogene <- merge(gogene, kegg, by=c("gene_id"), all=T)
head(gogene,2)
dim(gogene)
```


```{r}
geneInfo <- read.csv("../../output/WGCNA/WGCNA_ModuleMembership.csv") #this file was generated from the WGCNA analyses and contains the modules of interest
geneInfo<- plyr::rename(geneInfo, c("X"="gene_id"))
dim(geneInfo) # there are 9012 genes in our gene info file
```

```{r}
geneInfo <- merge(gogene, geneInfo, by=c("gene_id")) #merging the GO and Kegg info to module membership for the 9012 genes
```

Format GO terms to remove dashes and quotes and separate by semicolons (replace , with ;) in  GOs column
```{r}
geneInfo$GOs <- gsub(",", ";", geneInfo$GOs)
geneInfo$GOs <- gsub('"', "", geneInfo$GOs)
geneInfo$GOs <- gsub("-", NA, geneInfo$GOs)

geneInfo$KEGG_new[geneInfo$KEGG_new == ""] <- NA
```

```{r}
unique(geneInfo$moduleColor)
```

```{r}
geneInfo$Length<-lengths$transcript_lengths[match(geneInfo$gene_id, lengths$seqnames)]
dim(geneInfo)
write.csv(geneInfo, file = "../../output/WGCNA/GO_analysis/geneInfo_WGCNA.csv") #gene info for reference/supplement
```

## Modules

```{r}
calc_up_mods <- c("brown", "red", "black", "pink", "salmon", "blue")

nrow(geneInfo %>% dplyr::filter(moduleColor=="brown")) #942
nrow(geneInfo %>% dplyr::filter(moduleColor=="red")) #425
nrow(geneInfo %>% filter(moduleColor=="black")) #396
nrow(geneInfo %>% filter(moduleColor=="pink")) #220
nrow(geneInfo %>% filter(moduleColor=="salmon")) #154
nrow(geneInfo %>% filter(moduleColor=="blue")) #1989

sum(nrow(geneInfo %>% dplyr::filter(moduleColor=="brown")), nrow(geneInfo %>% dplyr::filter(moduleColor=="red")), nrow(geneInfo %>% filter(moduleColor=="black")), nrow(geneInfo %>% filter(moduleColor=="pink")), nrow(geneInfo %>% filter(moduleColor=="salmon")), nrow(geneInfo %>% filter(moduleColor=="blue")))

# 4126

calc_down_mods <- c("turquoise","magenta","lightcyan")

nrow(geneInfo %>% dplyr::filter(moduleColor=="turquoise")) #2558
nrow(geneInfo %>% dplyr::filter(moduleColor=="magenta")) #219
nrow(geneInfo %>% filter(moduleColor=="lightcyan")) #65

sum(nrow(geneInfo %>% dplyr::filter(moduleColor=="turquoise")), nrow(geneInfo %>% dplyr::filter(moduleColor=="magenta")), nrow(geneInfo %>% filter(moduleColor=="lightcyan")))

# 2842

other_mods <- c("green","yellow", "purple", "midnightblue","cyan","tan")

sum(nrow(geneInfo %>% dplyr::filter(moduleColor=="green")), nrow(geneInfo %>% dplyr::filter(moduleColor=="yellow")), nrow(geneInfo %>% filter(moduleColor=="purple")), nrow(geneInfo %>% filter(moduleColor=="midnightblue")), nrow(geneInfo %>% filter(moduleColor=="cyan")),nrow(geneInfo %>% filter(moduleColor=="tan")))

# 2044

# 4126 + 2842 + 2044 = 9012, which represents all of our genes
```

## Calcification - Up 

```{r}
### Generate vector with names of all genes 
ALL.vector <- c(geneInfo$gene_id)
```

```{r}
### Generate length vector for all genes 
LENGTH.vector <- as.integer(geneInfo$Length)
```

4126 genes are in the 6 modules significantly upregulated by calcification.

```{r}
### Generate vector with names in just the module we are analyzing
# ID.vector <- geneInfo %>%
#   filter(moduleColor==c("brown", "red", "black", "pink", "salmon", "green")) %>%
#   #get_rows(.data[[module]]))%>%
#   pull(gene_id)

ID.vector <- geneInfo %>%
  filter(moduleColor %in% c("brown", "red", "black", "pink", "salmon", "blue")) %>%
  pull(gene_id)

length(ID.vector) #4126

##Get a list of GO Terms for the all 9012 genes
GO.terms <- geneInfo %>%
  dplyr::select(gene_id, GOs) %>% rename(GOs = "GO.terms")

dim(GO.terms) #9012    2
```

```{r}
##Format to have one goterm per row with gene ID repeated
split <- strsplit(as.character(GO.terms$GO.terms), ";") 
split2 <- data.frame(v1 = rep.int(GO.terms$gene, sapply(split, length)), v2 = unlist(split)) 
colnames(split2) <- c("gene", "GO.terms")
GO.terms<-split2
```


```{r}
##Construct list of genes with 1 for genes in module and 0 for genes not in the module
gene.vector=as.integer(ALL.vector %in% ID.vector) 
names(gene.vector)<-ALL.vector#set names
#weight gene vector by bias for length of gene 
pwf<-nullp(gene.vector, ID.vector, bias.data=LENGTH.vector) 
```

### Run the GOSeq function by module color to test for GO term enrichment. Due to high number of enriched GO terms, I am using a p-value threshold of <0.05 and using `rrvgo` package to reduce redundancy in list of GO terms.   

```{r}
#run goseq using Wallenius method for all categories of GO terms 
GO.wall<-goseq(pwf, ID.vector, gene2cat=GO.terms, test.cats=c("GO:BP", "GO:MF", "GO:CC"), method="Wallenius", use_genes_without_cat=TRUE)
GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
colnames(GO)[1] <- "GOterm"
```

```{r}
#Filtering for p < 0.05
GO_05 <- GO %>%
        dplyr::filter(over_represented_pvalue<0.05) %>%
        dplyr::arrange(., ontology, over_represented_pvalue)
```

```{r}
#Write file of results 
write.csv(GO_05, file = "../../output/WGCNA/GO_analysis/goseq_pattern_calcification.csv")
```

```{r}
GO_05 <- read.csv("../../output/WGCNA/GO_analysis/goseq_pattern_calcification.csv")
```

```{r}
go_results_BP <- GO_05 %>%
      filter(ontology=="BP")%>%
      filter(over_represented_pvalue != "NA") %>%
      #filter(numInCat>10)%>%
      arrange(., over_represented_pvalue)
dim(go_results_BP)
head(go_results_BP)
```


Collapsing with simplifyEnrichment package

```{r}
library(simplifyEnrichment)
BP_terms <- go_results_BP$GOterm

Mat <- GO_similarity(BP_terms, ont = "BP", db = "org.Ce.eg.db")

#compare_clustering_methods(Mat, plot_type = "heatmap")

pdf(file = "../../output/WGCNA/GO_analysis/semantic_similarity/calc_up_mods_GOSEM.pdf", width = 7, height = 5)
simplifyGO(Mat, word_cloud_grob_param = list(max_width = 50), max_words=20)
dev.off()
```

```{r}
CalcUpMods_GO_P05 <- go_results_BP %>% 
  mutate(term = fct_reorder(term, over_represented_pvalue, .desc = TRUE)) %>%
  ggplot(aes(x=term, y=over_represented_pvalue) ) +
      geom_segment(aes(x=term ,xend=term, y=0, yend=over_represented_pvalue), color="grey") +
      geom_point(size=1, color="#69b3a2") +
      coord_flip() +
      theme(
        panel.grid.minor.y = element_blank(),
        panel.grid.major.y = element_blank(),
        legend.position="none"
      ) +
  xlab("") +
  ylab("over_represented_pvalue") +
  ggtitle("Biological Process GO Terms Enriched in Modules Upregulated in Calcification") + #add a main title
  theme(plot.title = element_text(face = 'bold',
                                  size = 3,
                                  hjust = 0)) +
  theme_bw() + #Set background color
  theme(panel.border = element_blank(), # Set border
                     panel.grid.major = element_blank(), #Set major gridlines
                     panel.grid.minor = element_blank(), #Set minor gridlines
                     axis.line = element_line(colour = "black"), #Set axes color
                    axis.text.y = element_text(size = 2), #set y axis labels size
        plot.background=element_blank(),#Set the plot background
        legend.position="none")

CalcUpMods_GO_P05
ggsave("../../output/WGCNA/GO_analysis/CalcUpMods_GO_P05.pdf", CalcUpMods_GO_P05, width = 12, height = 36, units = c("in"), dpi=300, limitsize=FALSE)
```


```{r}
library(rrvgo)
#Reduce/collapse GO term set with the rrvgo package 
simMatrix <- calculateSimMatrix(go_results_BP$GOterm,
                                orgdb="org.Ce.eg.db", #c. elegans database
                                ont="BP",
                                method="Rel")
```

```{r}
 #calculate similarity 
scores <- setNames(-log(go_results_BP$over_represented_pvalue), go_results_BP$GOterm)
reducedTerms <- reduceSimMatrix(simMatrix,
                                scores,
                                threshold=0.7,
                                orgdb="org.Ce.eg.db")
dim(reducedTerms)
```

```{r}
#keep only the goterms from the reduced list
go_results_BP_reduced <- go_results_BP %>%
  filter(GOterm %in% reducedTerms$go)
 #add in parent terms to list of go terms 
go_results_BP_reduced$ParentTerm <- reducedTerms$parentTerm[match(go_results_BP_reduced$GOterm, reducedTerms$go)]

length(unique(go_results_BP_reduced$ParentTerm))
```

The reduced list of terms is 508 terms that falls under 73 parent terms.

```{r}
write.csv(go_results_BP_reduced, "../../output/WGCNA/GO_analysis/goseq_pattern_calcification_filtered.csv")
```

Plot reduced terms
```{r}
#plot significantly enriched GO terms by Slim Category faceted by slim term 
 GO.plot_calcification <-  ggplot(go_results_BP_reduced, aes(x = ontology, y = term)) + 
    geom_point(aes(size=over_represented_pvalue)) + 
    scale_size(name="Over rep. p-value", trans="reverse", range=c(1,3))+
    facet_grid(ParentTerm ~ ., scales = "free", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
    strip.text.y = element_text(angle=0, size = 10),
    strip.text.x = element_text(size = 20),
    axis.text = element_text(size = 8),
    axis.title.x = element_blank(),
    axis.title.y = element_blank())
GO.plot_calcification

ggsave(filename="../../output/WGCNA/GO_analysis/GO.plot_calcification.pdf", plot=GO.plot_calcification, dpi=300, height=100, units="in", limitsize=FALSE)
```

```{r}
CalcUpMods_GO_P05_reduced <- go_results_BP_reduced %>% 
  mutate(term = fct_reorder(term, over_represented_pvalue, .desc = TRUE)) %>%
  ggplot(aes(x=term, y=over_represented_pvalue) ) +
      geom_segment(aes(x=term ,xend=term, y=0, yend=over_represented_pvalue), color="grey") +
      geom_point(size=1, color="#69b3a2") +
      coord_flip() +
      theme(
        panel.grid.minor.y = element_blank(),
        panel.grid.major.y = element_blank(),
        legend.position="none"
      ) +
  xlab("") +
  ylab("over_represented_pvalue") +
  ggtitle("Biological Process GO Terms Enriched in Modules Upregulated in Calcification") + #add a main title
  theme(plot.title = element_text(face = 'bold',
                                  size = 3,
                                  hjust = 0)) +
  theme_bw() + #Set background color
  theme(panel.border = element_blank(), # Set border
                     panel.grid.major = element_blank(), #Set major gridlines
                     panel.grid.minor = element_blank(), #Set minor gridlines
                     axis.line = element_line(colour = "black"), #Set axes color
                    axis.text.y = element_text(size = 2), #set y axis labels size
        plot.background=element_blank(),#Set the plot background
        legend.position="none")

CalcUpMods_GO_P05_reduced
ggsave("../../output/WGCNA/GO_analysis/CalcUpMods_GO_P05_reduced.pdf", CalcUpMods_GO_P05_reduced, width = 12, height = 36, units = c("in"), dpi=300, limitsize=FALSE)
```

### Calcification - Up - By Module

```{r}
### Generate vector with names of all genes 
ALL.vector <- c(geneInfo$gene_id)
```

```{r}
### Generate length vector for all genes 
LENGTH.vector <- as.integer(geneInfo$Length)
```


#### Brown

```{r}
ID.vector <- geneInfo %>%
  filter(moduleColor %in% c("brown")) %>%
  pull(gene_id)

length(ID.vector) #942

##Get a list of GO Terms for the all 9012 genes
GO.terms <- geneInfo %>%
  dplyr::select(gene_id, GOs) %>% rename(GOs = "GO.terms")

dim(GO.terms) #9012    2
```

```{r}
##Format to have one goterm per row with gene ID repeated
split <- strsplit(as.character(GO.terms$GO.terms), ";") 
split2 <- data.frame(v1 = rep.int(GO.terms$gene, sapply(split, length)), v2 = unlist(split)) 
colnames(split2) <- c("gene", "GO.terms")
GO.terms<-split2

##Construct list of genes with 1 for genes in module and 0 for genes not in the module
gene.vector=as.integer(ALL.vector %in% ID.vector) 
names(gene.vector)<-ALL.vector #set names
#weight gene vector by bias for length of gene 
pwf<-nullp(gene.vector, ID.vector, bias.data=LENGTH.vector) 

#run goseq using Wallenius method for all categories of GO terms 
GO.wall<-goseq(pwf, ID.vector, gene2cat=GO.terms, test.cats=c("GO:BP", "GO:MF", "GO:CC"), method="Wallenius", use_genes_without_cat=TRUE)
GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
colnames(GO)[1] <- "GOterm"
```

```{r}
#Filtering for p < 0.05
GO_05_brown <- GO %>%
        dplyr::filter(over_represented_pvalue<0.05) %>%
        dplyr::arrange(., ontology, over_represented_pvalue) %>% mutate(Module = "Brown")

go_results_BP <- GO_05_brown %>%
      filter(ontology=="BP")%>%
      filter(over_represented_pvalue != "NA") %>%
      #filter(numInCat>10)%>%
      arrange(., over_represented_pvalue)

dim(go_results_BP)
```

```{r}
#Write file of results 
write.csv(GO_05_brown, file = "../../output/WGCNA/GO_analysis/goseq_pattern_calcification_brown.csv")
```


```{r}
Top10Plot <- go_results_BP %>% 
    top_n(10, wt=-over_represented_pvalue) %>% 
    mutate(hitsPerc=numDEInCat/numInCat) %>% 
    ggplot(aes(x=hitsPerc, 
               y=reorder(term, -hitsPerc), 
               colour=over_represented_pvalue, 
               size=numDEInCat)) +
        geom_point() +
        scale_colour_gradient(low = "red", high = "blue", name = "p value",guide=guide_colorbar(reverse=TRUE)) + 
        expand_limits(x=0) +
        scale_y_discrete(labels = label_wrap(40)) + 
        labs(title = "Top 10 Enriched Biological Process GO Terms, Brown Module", x="Proportion of Genes in Module per GO Term", y="GO term", colour="p value", size="Count") +
        theme_bw() + 
        theme(axis.text.y = element_text(size = 8), 
              axis.title = element_text(size = 10), 
              plot.title = element_text(hjust = 0.5, size = 10)) 

Top10Plot

ggsave("../../output/WGCNA/GO_analysis/Top10_BP_by_mod/Brown.pdf", Top10Plot, width = 5, height = 5, units = c("in"), dpi=300, limitsize=FALSE)
```

Collapsing with simplifyEnrichment package

```{r}
library(simplifyEnrichment)
BP_terms <- go_results_BP$GOterm

Mat <- GO_similarity(BP_terms, ont = "BP", db = "org.Ce.eg.db")
pdf(file = "../../output/WGCNA/GO_analysis/semantic_similarity/calc_up_mods_GOSEM_brown.pdf", width = 7, height = 5)
simplifyGO(Mat, word_cloud_grob_param = list(max_width = 50),  min_term = 3, exclude_words = c("regulation", "process"), order_by_size = TRUE)
dev.off()
```

```{r}
library(rrvgo)
#Reduce/collapse GO term set with the rrvgo package 
simMatrix <- calculateSimMatrix(go_results_BP$GOterm,
                                orgdb="org.Ce.eg.db", #c. elegans database
                                ont="BP",
                                method="Rel")
```

```{r}
 #calculate similarity 
scores <- setNames(-log(go_results_BP$over_represented_pvalue), go_results_BP$GOterm)
reducedTerms <- reduceSimMatrix(simMatrix,
                                scores,
                                threshold=0.7,
                                orgdb="org.Ce.eg.db")
dim(reducedTerms)
```

```{r}
#keep only the goterms from the reduced list
go_results_BP_reduced <- go_results_BP %>%
  filter(GOterm %in% reducedTerms$go)
 #add in parent terms to list of go terms 
go_results_BP_reduced$ParentTerm <- reducedTerms$parentTerm[match(go_results_BP_reduced$GOterm, reducedTerms$go)]

length(unique(go_results_BP_reduced$ParentTerm))

go_results_BP_reduced <- go_results_BP_reduced %>% group_by(ParentTerm) %>% mutate("N_in_Parent" = n()) %>% ungroup()
```

The reduced list of terms is 157 terms that falls under 30 parent terms.

```{r}
write.csv(go_results_BP_reduced, "../../output/WGCNA/GO_analysis/goseq_brown_filtered.csv")
```

What about for p < 0.01?

```{r}
go_results_BP_reduced_01 <- go_results_BP_reduced %>% filter(over_represented_pvalue < 0.01) 
dim(go_results_BP_reduced_01)
length(unique(go_results_BP_reduced_01$ParentTerm))
```

This is only 41 terms that fall into 12 parent terms

```{r}
go_results_BP_reduced_Brown <- go_results_BP_reduced_01

top_10_parent <- go_results_BP_reduced_Brown %>% dplyr::select(ParentTerm, N_in_Parent) %>% distinct() %>% arrange(-N_in_Parent) %>% head(10)

Top10Plot <- go_results_BP_reduced_Brown %>% 
    filter(ParentTerm %in% top_10_parent$ParentTerm) %>%
    mutate(hitsPerc=numDEInCat/numInCat) %>% 
    arrange(-N_in_Parent) %>%
  mutate(ParentTerm = factor(ParentTerm, levels = unique(ParentTerm))) %>%
    ggplot(aes(x=numDEInCat, 
               y=reorder(term, numDEInCat), 
               colour=over_represented_pvalue)) +
        scale_colour_gradient(low = "red", high = "blue", name = "p value",guide=guide_colorbar(reverse=TRUE)) +
        expand_limits(x=1.3) +
        facet_grid(ParentTerm ~ .,scales="free_y",labeller = labeller(ParentTerm = label_wrap_gen(width = 25))) + 
        geom_segment(aes(y=term ,yend=term, x=0, xend=numDEInCat), color="black",linewidth=0.3) +
       geom_point() +
    scale_y_discrete(labels = label_wrap(30)) + 
        labs(title = "Enriched Biological Process GO Terms p < 0.01,\nFacet by Parent Term (Top 10), Brown Module", x="Count", y="GO term", colour="p-value") +
        theme_bw() + 
        theme(axis.text.y = element_text(size = 4), 
              axis.title = element_text(size = 10), 
              plot.title = element_text(hjust = 0.5, size = 10),
              strip.text.y = element_text(size = 6,angle = 0,face = "bold"))

Top10Plot

ggsave("../../output/WGCNA/GO_analysis/Top10_BP_by_mod/Brown_allterms_top10parent_filtp01.pdf", Top10Plot, width = 5, height = 7, units = c("in"), dpi=300, limitsize=FALSE)

top_5_parent <- go_results_BP_reduced_Brown %>% dplyr::select(ParentTerm, N_in_Parent) %>% distinct() %>% arrange(-N_in_Parent) %>% head(5)

Top5Plot <- go_results_BP_reduced_Brown %>% 
    filter(ParentTerm %in% top_5_parent$ParentTerm) %>%
    filter(over_represented_pvalue <0.01) %>%
    mutate(hitsPerc=numDEInCat/numInCat) %>% 
    arrange(-N_in_Parent) %>%
  mutate(ParentTerm = factor(ParentTerm, levels = unique(ParentTerm))) %>%
    
    ggplot(aes(x=numDEInCat, 
               y=reorder(term, numDEInCat), 
               colour=over_represented_pvalue)) +
        scale_colour_gradient(low = "red", high = "blue", name = "p value",guide=guide_colorbar(reverse=TRUE)) +
        expand_limits(x=1.3) +
        facet_grid(ParentTerm ~ .,scales="free_y", space = "free",labeller = labeller(ParentTerm = label_wrap_gen(width = 25))) + 
        geom_segment(aes(y=term ,yend=term, x=0, xend=numDEInCat), color="black",linewidth=0.3) +
       geom_point() +
    scale_y_discrete(labels = label_wrap(30)) + 
        labs(title = "Enriched Biological Process GO Terms p < 0.01,\nFacet by Parent Term (Top 5), Brown Module", x="Count", y="GO term", colour="p-value") +
        theme_bw() + 
        theme(axis.text.y = element_text(size = 4), 
              axis.title = element_text(size = 10), 
              plot.title = element_text(hjust = 0.5, size = 10),
              strip.text.y = element_text(size = 6,angle = 0,face = "bold"))

Top5Plot

ggsave("../../output/WGCNA/GO_analysis/Top10_BP_by_mod/Brown_allterms_top5parent_filtp01.pdf", Top5Plot, width = 5, height = 5, units = c("in"), dpi=300, limitsize=FALSE)
```


#### Red

```{r}
ID.vector <- geneInfo %>%
  filter(moduleColor %in% c("red")) %>%
  pull(gene_id)

length(ID.vector) #425

##Get a list of GO Terms for the all 9012 genes
GO.terms <- geneInfo %>%
  dplyr::select(gene_id, GOs) %>% rename(GOs = "GO.terms")

dim(GO.terms) #9012    2
```

```{r}
##Format to have one goterm per row with gene ID repeated
split <- strsplit(as.character(GO.terms$GO.terms), ";") 
split2 <- data.frame(v1 = rep.int(GO.terms$gene, sapply(split, length)), v2 = unlist(split)) 
colnames(split2) <- c("gene", "GO.terms")
GO.terms<-split2

##Construct list of genes with 1 for genes in module and 0 for genes not in the module
gene.vector=as.integer(ALL.vector %in% ID.vector) 
names(gene.vector)<-ALL.vector#set names
#weight gene vector by bias for length of gene 
pwf<-nullp(gene.vector, ID.vector, bias.data=LENGTH.vector) 

#run goseq using Wallenius method for all categories of GO terms 
GO.wall<-goseq(pwf, ID.vector, gene2cat=GO.terms, test.cats=c("GO:BP", "GO:MF", "GO:CC"), method="Wallenius", use_genes_without_cat=TRUE)
GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
colnames(GO)[1] <- "GOterm"
```

```{r}
#Filtering for p < 0.05
GO_05_red <- GO %>%
        dplyr::filter(over_represented_pvalue<0.05) %>%
        dplyr::arrange(., ontology, over_represented_pvalue) %>% mutate(Module = "Red")

go_results_BP <- GO_05_red %>%
      filter(ontology=="BP")%>%
      filter(over_represented_pvalue != "NA") %>%
      #filter(numInCat>10)%>%
      arrange(., over_represented_pvalue)

dim(go_results_BP)
```

```{r}
#Write file of results 
write.csv(GO_05_red, file = "../../output/WGCNA/GO_analysis/goseq_pattern_calcification_red.csv")
```


```{r}
Top10Plot <- go_results_BP %>% 
    top_n(10, wt=-over_represented_pvalue) %>% 
    mutate(hitsPerc=numDEInCat/numInCat) %>% 
    ggplot(aes(x=hitsPerc, 
               y=reorder(term, -hitsPerc), 
               colour=over_represented_pvalue, 
               size=numDEInCat)) +
        geom_point() +
        scale_colour_gradient(low = "red", high = "blue", name = "p value",guide=guide_colorbar(reverse=TRUE)) + 
  
        expand_limits(x=0) +
        scale_y_discrete(labels = label_wrap(40)) + 
        labs(title = "Top 10 Enriched Biological Process GO Terms, Red Module", x="Proportion of Genes in Module per GO Term", y="GO term", colour="p value", size="Count") +
        theme_bw() + 
        theme(axis.text.y = element_text(size = 8), 
              axis.title = element_text(size = 10), 
              plot.title = element_text(hjust = 0.5, size = 10)) 

Top10Plot

ggsave("../../output/WGCNA/GO_analysis/Top10_BP_by_mod/Red.pdf", Top10Plot, width = 5, height = 5, units = c("in"), dpi=300, limitsize=FALSE)
```


Collapsing with simplifyEnrichment package

```{r}
library(simplifyEnrichment)
BP_terms <- go_results_BP$GOterm

Mat <- GO_similarity(BP_terms, ont = "BP", db = "org.Ce.eg.db")
pdf(file = "../../output/WGCNA/GO_analysis/semantic_similarity/calc_up_mods_GOSEM_red.pdf", width = 7, height = 5)
result <-  simplifyGO(Mat, word_cloud_grob_param = list(max_width = 50),  min_term = 3, exclude_words = c("regulation", "process"), order_by_size = TRUE)
result
dev.off()

# Following filtering of small clusters is based on the simplifyGO/ht_clusters source code for merging the small clusters together

# Filter clusters smaller than 3
cluster_sizes <- table(result$cluster) # table function makes a table with the size of each cluster (freqeuncy) 
large_clusters <- as.numeric(names(cluster_sizes)[cluster_sizes >= 3]) # Identify clusters with frequenct/size less than 3
filtered_indices <- result$cluster %in% large_clusters # Filter out the indices belonging to small clusters
filtered_Mat <- Mat[filtered_indices, filtered_indices] # Create a new matrix with only the indicies from larger clusters

pdf(file = "../../output/WGCNA/GO_analysis/semantic_similarity/calc_up_mods_GOSEM_red_filtered.pdf", width = 7, height = 5)
simplifyGO(filtered_Mat, word_cloud_grob_param = list(max_width = 50),  min_term = 1, exclude_words = c("regulation", "process"), order_by_size = TRUE)
dev.off()
```


```{r}
library(rrvgo)
#Reduce/collapse GO term set with the rrvgo package 
simMatrix <- calculateSimMatrix(go_results_BP$GOterm,
                                orgdb="org.Ce.eg.db", #c. elegans database
                                ont="BP",
                                method="Rel")
```

```{r}
 #calculate similarity 
scores <- setNames(-log(go_results_BP$over_represented_pvalue), go_results_BP$GOterm)
reducedTerms <- reduceSimMatrix(simMatrix,
                                scores,
                                threshold=0.7,
                                orgdb="org.Ce.eg.db")
dim(reducedTerms)
```

```{r}
#keep only the goterms from the reduced list
go_results_BP_reduced <- go_results_BP %>%
  filter(GOterm %in% reducedTerms$go)
 #add in parent terms to list of go terms 
go_results_BP_reduced$ParentTerm <- reducedTerms$parentTerm[match(go_results_BP_reduced$GOterm, reducedTerms$go)]

length(unique(go_results_BP_reduced$ParentTerm))

go_results_BP_reduced <- go_results_BP_reduced %>% group_by(ParentTerm) %>% mutate("N_in_Parent" = n()) %>% ungroup()
```

The reduced list of terms is 85 terms that falls under 26 parent terms.

```{r}
write.csv(go_results_BP_reduced, "../../output/WGCNA/GO_analysis/goseq_red_filtered.csv")
```


What about for p < 0.01?

```{r}
go_results_BP_reduced_01 <- go_results_BP_reduced %>% filter(over_represented_pvalue < 0.01) 
dim(go_results_BP_reduced_01)
length(unique(go_results_BP_reduced_01$ParentTerm))
```

This is only 16 terms that fall into 9 parent terms

```{r}
go_results_BP_reduced_Red <- go_results_BP_reduced_01

top_10_parent <- go_results_BP_reduced_Red %>% dplyr::select(ParentTerm, N_in_Parent) %>% distinct() %>% arrange(-N_in_Parent) %>% head(10)

Top10Plot <- go_results_BP_reduced_Red %>% 
    filter(ParentTerm %in% top_10_parent$ParentTerm) %>%
    mutate(hitsPerc=numDEInCat/numInCat) %>% 
    arrange(-N_in_Parent) %>%
  mutate(ParentTerm = factor(ParentTerm, levels = unique(ParentTerm))) %>%
    ggplot(aes(x=numDEInCat, 
               y=reorder(term, numDEInCat), 
               colour=over_represented_pvalue)) +
        scale_colour_gradient(low = "red", high = "blue", name = "p value",guide=guide_colorbar(reverse=TRUE)) +
        expand_limits(x=1.3) +
        facet_grid(ParentTerm ~ .,scales="free_y",labeller = labeller(ParentTerm = label_wrap_gen(width = 25))) + 
        geom_segment(aes(y=term ,yend=term, x=0, xend=numDEInCat), color="black",linewidth=0.3) +
       geom_point() +
    scale_y_discrete(labels = label_wrap(30)) + 
        labs(title = "Enriched Biological Process GO Terms p < 0.01,\nFacet by Parent Term, Red Module", x="Count", y="GO term", colour="p-value") +
        theme_bw() + 
        theme(axis.text.y = element_text(size = 4), 
              axis.title = element_text(size = 10), 
              plot.title = element_text(hjust = 0.5, size = 10),
              strip.text.y = element_text(size = 6,angle = 0,face = "bold"))

Top10Plot

ggsave("../../output/WGCNA/GO_analysis/Top10_BP_by_mod/Red_allterms_top10parent_filtp01.pdf", Top10Plot, width = 5, height = 7, units = c("in"), dpi=300, limitsize=FALSE)

top_5_parent <- go_results_BP_reduced_Red %>% dplyr::select(ParentTerm, N_in_Parent) %>% distinct() %>% arrange(-N_in_Parent) %>% head(5)

Top5Plot <- go_results_BP_reduced_Red %>% 
    filter(ParentTerm %in% top_5_parent$ParentTerm) %>%
    filter(over_represented_pvalue <0.01) %>%
    mutate(hitsPerc=numDEInCat/numInCat) %>% 
    arrange(-N_in_Parent) %>%
  mutate(ParentTerm = factor(ParentTerm, levels = unique(ParentTerm))) %>%
    
    ggplot(aes(x=numDEInCat, 
               y=reorder(term, numDEInCat), 
               colour=over_represented_pvalue)) +
        scale_colour_gradient(low = "red", high = "blue", name = "p value",guide=guide_colorbar(reverse=TRUE)) +
        expand_limits(x=1.3) +
        facet_grid(ParentTerm ~ .,scales="free_y", space = "free",labeller = labeller(ParentTerm = label_wrap_gen(width = 25))) + 
        geom_segment(aes(y=term ,yend=term, x=0, xend=numDEInCat), color="black",linewidth=0.3) +
       geom_point() +
    scale_y_discrete(labels = label_wrap(30)) + 
        labs(title = "Enriched Biological Process GO Terms p < 0.01,\nFacet by Parent Term (Top 5), Red Module", x="Count", y="GO term", colour="p-value") +
        theme_bw() + 
        theme(axis.text.y = element_text(size = 4), 
              axis.title = element_text(size = 10), 
              plot.title = element_text(hjust = 0.5, size = 10),
              strip.text.y = element_text(size = 6,angle = 0,face = "bold"))

Top5Plot

ggsave("../../output/WGCNA/GO_analysis/Top10_BP_by_mod/Red_allterms_top5parent_filtp01.pdf", Top5Plot, width = 5, height = 5, units = c("in"), dpi=300, limitsize=FALSE)
```

#### Black

```{r}
ID.vector <- geneInfo %>%
  filter(moduleColor %in% c("black")) %>%
  pull(gene_id)

length(ID.vector) #396

##Get a list of GO Terms for the all 9012 genes
GO.terms <- geneInfo %>%
  dplyr::select(gene_id, GOs) %>% rename(GOs = "GO.terms")

dim(GO.terms) #9012    2
```

```{r}
##Format to have one goterm per row with gene ID repeated
split <- strsplit(as.character(GO.terms$GO.terms), ";") 
split2 <- data.frame(v1 = rep.int(GO.terms$gene, sapply(split, length)), v2 = unlist(split)) 
colnames(split2) <- c("gene", "GO.terms")
GO.terms<-split2

##Construct list of genes with 1 for genes in module and 0 for genes not in the module
gene.vector=as.integer(ALL.vector %in% ID.vector) 
names(gene.vector)<-ALL.vector#set names
#weight gene vector by bias for length of gene 
pwf<-nullp(gene.vector, ID.vector, bias.data=LENGTH.vector) 

#run goseq using Wallenius method for all categories of GO terms 
GO.wall<-goseq(pwf, ID.vector, gene2cat=GO.terms, test.cats=c("GO:BP", "GO:MF", "GO:CC"), method="Wallenius", use_genes_without_cat=TRUE)
GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
colnames(GO)[1] <- "GOterm"
```

```{r}
#Filtering for p < 0.05
GO_05_black <- GO %>%
        dplyr::filter(over_represented_pvalue<0.05) %>%
        dplyr::arrange(., ontology, over_represented_pvalue) %>% mutate(Module = "Black")

go_results_BP <- GO_05_black %>%
      filter(ontology=="BP")%>%
      filter(over_represented_pvalue != "NA") %>%
      #filter(numInCat>10)%>%
      arrange(., over_represented_pvalue)

dim(go_results_BP)
```

```{r}
Top10Plot <- go_results_BP %>% 
    top_n(10, wt=-over_represented_pvalue) %>% 
    mutate(hitsPerc=numDEInCat/numInCat) %>% 
    ggplot(aes(x=hitsPerc, 
               y=reorder(term, -hitsPerc), 
               colour=over_represented_pvalue, 
               size=numDEInCat)) +
        geom_point() +
        scale_colour_gradient(low = "red", high = "blue", name = "p value",guide=guide_colorbar(reverse=TRUE)) + 
        expand_limits(x=0) +
        scale_y_discrete(labels = label_wrap(40)) + 
        labs(title = "Top 10 Enriched Biological Process GO Terms, Black Module", x="Proportion of Genes in Module per GO Term", y="GO term", colour="p value", size="Count") +
        theme_bw() + 
        theme(axis.text.y = element_text(size = 8), 
              axis.title = element_text(size = 10), 
              plot.title = element_text(hjust = 0.5, size = 10)) 

Top10Plot

ggsave("../../output/WGCNA/GO_analysis/Top10_BP_by_mod/Black.pdf", Top10Plot, width = 5, height = 5, units = c("in"), dpi=300, limitsize=FALSE)
```

Collapsing with simplifyEnrichment package

```{r}
library(simplifyEnrichment)
BP_terms <- go_results_BP$GOterm

Mat <- GO_similarity(BP_terms, ont = "BP", db = "org.Ce.eg.db")
pdf(file = "../../output/WGCNA/GO_analysis/semantic_similarity/calc_up_mods_GOSEM_black.pdf", width = 7, height = 5)
simplifyGO(Mat, word_cloud_grob_param = list(max_width = 50), min_term = 3, exclude_words = c("regulation", "process"), order_by_size = TRUE)
dev.off()
```

#### Pink

```{r}
ID.vector <- geneInfo %>%
  filter(moduleColor %in% c("pink")) %>%
  pull(gene_id)

length(ID.vector) #220

##Get a list of GO Terms for the all 9012 genes
GO.terms <- geneInfo %>%
  dplyr::select(gene_id, GOs) %>% rename(GOs = "GO.terms")

dim(GO.terms) #9012    2
```

```{r}
##Format to have one goterm per row with gene ID repeated
split <- strsplit(as.character(GO.terms$GO.terms), ";") 
split2 <- data.frame(v1 = rep.int(GO.terms$gene, sapply(split, length)), v2 = unlist(split)) 
colnames(split2) <- c("gene", "GO.terms")
GO.terms<-split2

##Construct list of genes with 1 for genes in module and 0 for genes not in the module
gene.vector=as.integer(ALL.vector %in% ID.vector) 
names(gene.vector)<-ALL.vector#set names
#weight gene vector by bias for length of gene 
pwf<-nullp(gene.vector, ID.vector, bias.data=LENGTH.vector) 

#run goseq using Wallenius method for all categories of GO terms 
GO.wall<-goseq(pwf, ID.vector, gene2cat=GO.terms, test.cats=c("GO:BP", "GO:MF", "GO:CC"), method="Wallenius", use_genes_without_cat=TRUE)
GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
colnames(GO)[1] <- "GOterm"
```

```{r}
#Filtering for p < 0.05
GO_05_pink <- GO %>%
        dplyr::filter(over_represented_pvalue<0.05) %>%
        dplyr::arrange(., ontology, over_represented_pvalue) %>% mutate(Module = "Pink")

go_results_BP <- GO_05_pink %>%
      filter(ontology=="BP")%>%
      filter(over_represented_pvalue != "NA") %>%
      #filter(numInCat>10)%>%
      arrange(., over_represented_pvalue)

dim(go_results_BP)
```

```{r}
Top10Plot <- go_results_BP %>% 
    top_n(10, wt=-over_represented_pvalue) %>% 
    mutate(hitsPerc=numDEInCat/numInCat) %>% 
    ggplot(aes(x=hitsPerc, 
               y=reorder(term, -hitsPerc), 
               colour=over_represented_pvalue, 
               size=numDEInCat)) +
        geom_point() +
        scale_colour_gradient(low = "red", high = "blue", name = "p value",guide=guide_colorbar(reverse=TRUE)) + 
        expand_limits(x=0) +
        scale_y_discrete(labels = label_wrap(40)) + 
        labs(title = "Top 10 Enriched Biological Process GO Terms, Pink Module", x="Proportion of Genes in Module per GO Term", y="GO term", colour="p value", size="Count") +
        theme_bw() +
        theme(axis.text.y = element_text(size = 8), 
              axis.title = element_text(size = 10), 
              plot.title = element_text(hjust = 0.5, size = 10)) 

Top10Plot

ggsave("../../output/WGCNA/GO_analysis/Top10_BP_by_mod/Pink.pdf", Top10Plot, width = 5, height = 5, units = c("in"), dpi=300, limitsize=FALSE)
```

Collapsing with simplifyEnrichment package

```{r}
library(simplifyEnrichment)
BP_terms <- go_results_BP$GOterm

Mat <- GO_similarity(BP_terms, ont = "BP", db = "org.Ce.eg.db")
pdf(file = "../../output/WGCNA/GO_analysis/semantic_similarity/calc_up_mods_GOSEM_pink.pdf", width = 7, height = 5)
simplifyGO(Mat, word_cloud_grob_param = list(max_width = 50), min_term = 3, exclude_words = c("regulation", "process"), order_by_size = TRUE)
dev.off()
```

#### Salmon

```{r}
ID.vector <- geneInfo %>%
  filter(moduleColor %in% c("salmon")) %>%
  pull(gene_id)

length(ID.vector) #154

##Get a list of GO Terms for the all 9012 genes
GO.terms <- geneInfo %>%
  dplyr::select(gene_id, GOs) %>% rename(GOs = "GO.terms")

dim(GO.terms) #9012    2
```

```{r}
##Format to have one goterm per row with gene ID repeated
split <- strsplit(as.character(GO.terms$GO.terms), ";") 
split2 <- data.frame(v1 = rep.int(GO.terms$gene, sapply(split, length)), v2 = unlist(split)) 
colnames(split2) <- c("gene", "GO.terms")
GO.terms<-split2

##Construct list of genes with 1 for genes in module and 0 for genes not in the module
gene.vector=as.integer(ALL.vector %in% ID.vector) 
names(gene.vector)<-ALL.vector#set names
#weight gene vector by bias for length of gene 
pwf<-nullp(gene.vector, ID.vector, bias.data=LENGTH.vector) 

#run goseq using Wallenius method for all categories of GO terms 
GO.wall<-goseq(pwf, ID.vector, gene2cat=GO.terms, test.cats=c("GO:BP", "GO:MF", "GO:CC"), method="Wallenius", use_genes_without_cat=TRUE)
GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
colnames(GO)[1] <- "GOterm"
```

```{r}
#Filtering for p < 0.05
GO_05_salmon <- GO %>%
        dplyr::filter(over_represented_pvalue<0.05) %>%
        dplyr::arrange(., ontology, over_represented_pvalue) %>% mutate(Module = "Salmon")

go_results_BP <- GO_05_salmon %>%
      filter(ontology=="BP")%>%
      filter(over_represented_pvalue != "NA") %>%
      #filter(numInCat>10)%>%
      arrange(., over_represented_pvalue)

dim(go_results_BP)
```

```{r}
Top10Plot <- go_results_BP %>% 
    top_n(10, wt=-over_represented_pvalue) %>% 
    mutate(hitsPerc=numDEInCat/numInCat) %>% 
    ggplot(aes(x=hitsPerc, 
               y=reorder(term, -hitsPerc), 
               colour=over_represented_pvalue, 
               size=numDEInCat)) +
        geom_point() +
        scale_colour_gradient(low = "red", high = "blue", name = "p value",guide=guide_colorbar(reverse=TRUE)) + 
        expand_limits(x=0) +
        scale_y_discrete(labels = label_wrap(40)) + 
        labs(title = "Top 10 Enriched Biological Process GO Terms, Salmon Module", x="Proportion of Genes in Module per GO Term", y="GO term", colour="p value", size="Count") +
        theme_bw() + 
        theme(axis.text.y = element_text(size = 8), 
              axis.title = element_text(size = 10), 
              plot.title = element_text(hjust = 0.5, size = 10)) 

Top10Plot

ggsave("../../output/WGCNA/GO_analysis/Top10_BP_by_mod/Salmon.pdf", Top10Plot, width = 5, height = 5, units = c("in"), dpi=300, limitsize=FALSE)
```

Collapsing with simplifyEnrichment package

```{r}
library(simplifyEnrichment)
BP_terms <- go_results_BP$GOterm

Mat <- GO_similarity(BP_terms, ont = "BP", db = "org.Ce.eg.db")
pdf(file = "../../output/WGCNA/GO_analysis/semantic_similarity/calc_up_mods_GOSEM_salmon.pdf", width = 7, height = 5)
simplifyGO(Mat, word_cloud_grob_param = list(max_width = 50), min_term = 3, exclude_words = c("regulation", "process"), order_by_size = TRUE)
dev.off()
```

#### Blue

```{r}
ID.vector <- geneInfo %>%
  filter(moduleColor %in% c("blue")) %>%
  pull(gene_id)

length(ID.vector) #1989

##Get a list of GO Terms for the all 9012 genes
GO.terms <- geneInfo %>%
  dplyr::select(gene_id, GOs) %>% rename(GOs = "GO.terms")

dim(GO.terms) #9012    2
```

```{r}
##Format to have one goterm per row with gene ID repeated
split <- strsplit(as.character(GO.terms$GO.terms), ";") 
split2 <- data.frame(v1 = rep.int(GO.terms$gene, sapply(split, length)), v2 = unlist(split)) 
colnames(split2) <- c("gene", "GO.terms")
GO.terms<-split2

##Construct list of genes with 1 for genes in module and 0 for genes not in the module
gene.vector=as.integer(ALL.vector %in% ID.vector) 
names(gene.vector)<-ALL.vector#set names
#weight gene vector by bias for length of gene 
pwf<-nullp(gene.vector, ID.vector, bias.data=LENGTH.vector) 

#run goseq using Wallenius method for all categories of GO terms 
GO.wall<-goseq(pwf, ID.vector, gene2cat=GO.terms, test.cats=c("GO:BP", "GO:MF", "GO:CC"), method="Wallenius", use_genes_without_cat=TRUE)
GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
colnames(GO)[1] <- "GOterm"
```

```{r}
#Filtering for p < 0.05
GO_05_blue <- GO %>%
        dplyr::filter(over_represented_pvalue<0.05) %>%
        dplyr::arrange(., ontology, over_represented_pvalue) %>% mutate(Module = "Blue")

go_results_BP <- GO_05_blue %>%
      filter(ontology=="BP")%>%
      filter(over_represented_pvalue != "NA") %>%
      #filter(numInCat>10)%>%
      arrange(., over_represented_pvalue)

dim(go_results_BP)
```

```{r}
Top10Plot <- go_results_BP %>% 
    top_n(10, wt=-over_represented_pvalue) %>% 
    mutate(hitsPerc=numDEInCat/numInCat) %>% 
    ggplot(aes(x=hitsPerc, 
               y=reorder(term, -hitsPerc), 
               colour=over_represented_pvalue, 
               size=numDEInCat)) +
        geom_point() +
        scale_colour_gradient(low = "red", high = "blue", name = "p value",guide=guide_colorbar(reverse=TRUE)) + 
        expand_limits(x=0) +
        scale_y_discrete(labels = label_wrap(40)) + 
        labs(title = "Top 10 Enriched Biological Process GO Terms, Blue Module", x="Proportion of Genes in Module per GO Term", y="GO term", colour="p value", size="Count") +
        theme_bw() + 
        theme(axis.text.y = element_text(size = 8), 
              axis.title = element_text(size = 10), 
              plot.title = element_text(hjust = 0.5, size = 10)) 

Top10Plot

ggsave("../../output/WGCNA/GO_analysis/Top10_BP_by_mod/Blue.pdf", Top10Plot, width = 5, height = 5, units = c("in"), dpi=300, limitsize=FALSE)
```

Collapsing with simplifyEnrichment package

```{r}
library(simplifyEnrichment)
BP_terms <- go_results_BP$GOterm

Mat <- GO_similarity(BP_terms, ont = "BP", db = "org.Ce.eg.db")
pdf(file = "../../output/WGCNA/GO_analysis/semantic_similarity/calc_up_mods_GOSEM_blue.pdf", width = 7, height = 5)
simplifyGO(Mat, word_cloud_grob_param = list(max_width = 50))
dev.off()
```

### Combining all the individual module lists for Calcification - Up

```{r}
GO_05_UP <- rbind(GO_05_brown,GO_05_red,GO_05_black,GO_05_pink,GO_05_salmon,GO_05_blue)

length(GO_05_UP$GOterm) #3810 enriched GO terms
length(unique(GO_05_UP$GOterm)) #3617 enriched GO terms

# Collapse GO terms that are duplicated accross modules - sort by GO term and p-value, then keep the row with the lowest p-value, collapsing the module names so you know which modules were represented

# Identify and collapse duplicates
GO_05_UP <- GO_05_UP %>%
  arrange(GOterm, over_represented_pvalue) %>%
  group_by(GOterm) %>% mutate(Module = paste(unique(Module), collapse = ";")) %>%
  group_by(GOterm) %>%
  slice_min(order_by = over_represented_pvalue, n = 1) %>%
  ungroup()

length(GO_05_UP$GOterm) #3617 enriched GO terms
length(unique(GO_05_UP$GOterm)) #3617 enriched GO terms, all are unique

#Write file of results 
write.csv(GO_05_UP, file = "../../output/WGCNA/GO_analysis/goseq_pattern_calcification_up_bymod.csv")
```

## Calcification - Down

2842 genes are in the 4 modules significantly downregulated by calcification.

```{r}
### Generate vector with names of all genes 
ALL.vector <- c(geneInfo$gene_id)
```

```{r}
### Generate length vector for all genes 
LENGTH.vector <- as.integer(geneInfo$Length)
```

```{r}
### Generate vector with names in just the module we are analyzing
ID.vector <- geneInfo%>%
  filter(moduleColor %in% c("turquoise","magenta","lightcyan"))%>%
  pull(gene_id)

length(ID.vector) #2842

##Get a list of GO Terms for the all 9012 genes
GO.terms <- geneInfo %>%
  dplyr::select(gene_id, GOs) %>% rename(GOs = "GO.terms")

dim(GO.terms) #9012    2
```

```{r}
##Format to have one goterm per row with gene ID repeated
split <- strsplit(as.character(GO.terms$GO.terms), ";") 
split2 <- data.frame(v1 = rep.int(GO.terms$gene, sapply(split, length)), v2 = unlist(split)) 
colnames(split2) <- c("gene", "GO.terms")
GO.terms<-split2
```


```{r}
##Construct list of genes with 1 for genes in module and 0 for genes not in the module
gene.vector=as.integer(ALL.vector %in% ID.vector) 
names(gene.vector)<-ALL.vector#set names
#weight gene vector by bias for length of gene 
pwf<-nullp(gene.vector, ID.vector, bias.data=LENGTH.vector) 
```
### Run the GOSeq function by module color to test for GO term enrichment. Due to high number of enriched GO terms, I am using a p-value threshold of <0.05 and using `rrvgo` package to reduce redundancy in list of GO terms.   

```{r}
#run goseq using Wallenius method for all categories of GO terms 
GO.wall<-goseq(pwf, ID.vector, gene2cat=GO.terms, test.cats=c("GO:BP", "GO:MF", "GO:CC"), method="Wallenius", use_genes_without_cat=TRUE)
GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
colnames(GO)[1] <- "GOterm"
```

```{r}
#Filtering for p < 0.05
GO_05 <- GO %>%
        dplyr::filter(over_represented_pvalue<0.05) %>%
        dplyr::arrange(., ontology, over_represented_pvalue)
```

```{r}
#Write file of results 
write.csv(GO_05, file = "../../output/WGCNA/GO_analysis/goseq_pattern_calcification_down.csv")
```

```{r}
GO_05 <-read.csv("../../output/WGCNA/GO_analysis/goseq_pattern_calcification_down.csv")
```

```{r}
go_results_BP <- GO_05 %>%
      filter(ontology=="BP") %>%
      filter(over_represented_pvalue != "NA") %>%
      #filter(numInCat > 10) %>%
      arrange(., over_represented_pvalue)

dim(go_results_BP)
head(go_results_BP)
```


Collapsing with simplifyEnrichment package

```{r}
library(simplifyEnrichment)
BP_terms <- go_results_BP$GOterm

Mat <- GO_similarity(BP_terms, ont = "BP", db = "org.Ce.eg.db")
pdf(file = "../../output/WGCNA/GO_analysis/semantic_similarity/calc_down_mods_GOSEM.pdf", width = 7, height = 5)
simplifyGO(Mat, word_cloud_grob_param = list(max_width = 50), max_words=20)
dev.off()
```

```{r}
CalcDownMods_GO_P05 <- go_results_BP %>% 
  mutate(term = fct_reorder(term, over_represented_pvalue, .desc = TRUE)) %>%
  ggplot(aes(x=term, y=over_represented_pvalue) ) +
      geom_segment(aes(x=term ,xend=term, y=0, yend=over_represented_pvalue), color="grey") +
      geom_point(size=1, color="#69b3a2") +
      coord_flip() +
      theme(
        panel.grid.minor.y = element_blank(),
        panel.grid.major.y = element_blank(),
        legend.position="none"
      ) +
  xlab("") +
  ylab("over_represented_pvalue") +
  ggtitle("Biological Process GO Terms Enriched in Modules Downregulated in Calcification") + #add a main title
  theme(plot.title = element_text(face = 'bold',
                                  size = 3,
                                  hjust = 0)) +
  theme_bw() + #Set background color
  theme(panel.border = element_blank(), # Set border
                     panel.grid.major = element_blank(), #Set major gridlines
                     panel.grid.minor = element_blank(), #Set minor gridlines
                     axis.line = element_line(colour = "black"), #Set axes color
                    axis.text.y = element_text(size = 2), #set y axis labels size
        plot.background=element_blank(),#Set the plot background
        legend.position="none")

CalcDownMods_GO_P05
ggsave("../../output/WGCNA/GO_analysis/CalcDownMods_GO_P05.pdf", CalcDownMods_GO_P05, width = 12, height = 36, units = c("in"), dpi=300, limitsize=FALSE)
```

```{r}
library(rrvgo)
#Reduce/collapse GO term set with the rrvgo package 
simMatrix <- calculateSimMatrix(go_results_BP$GOterm,
                                orgdb="org.Ce.eg.db", #c. elegans database
                                ont="BP",
                                method="Rel")
```

```{r}
 #calculate similarity 
scores <- setNames(-log(go_results_BP$over_represented_pvalue), go_results_BP$GOterm)
reducedTerms <- reduceSimMatrix(simMatrix,
                                scores,
                                threshold=0.7,
                                orgdb="org.Ce.eg.db")
dim(reducedTerms)
```

```{r}
#keep only the goterms from the reduced list
go_results_BP_reduced <- go_results_BP %>%
  filter(GOterm %in% reducedTerms$go)
 #add in parent terms to list of go terms 
go_results_BP_reduced$ParentTerm <- reducedTerms$parentTerm[match(go_results_BP_reduced$GOterm, reducedTerms$go)]

length(unique(go_results_BP_reduced$ParentTerm))
```

The reduced list of terms is 226 terms that falls under 38 parent terms.

```{r}
write.csv(go_results_BP_reduced, "../../output/WGCNA/GO_analysis/goseq_pattern_calcification_down_filtered.csv")
```


```{r}
#plot significantly enriched GO terms by Slim Category faceted by slim term 
 GO.plot_calcification_down <-  ggplot(go_results_BP_reduced, aes(x = ontology, y = term)) + 
    geom_point(aes(size=over_represented_pvalue)) + 
    scale_size(name="Over rep. p-value", trans="reverse", range=c(1,3))+
    facet_grid(ParentTerm ~ ., scales = "free", labeller = label_wrap_gen(width = 5, multi_line = TRUE))+
    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
    strip.text.y = element_text(angle=0, size = 10),
    strip.text.x = element_text(size = 20),
    axis.text = element_text(size = 8),
    axis.title.x = element_blank(),
    axis.title.y = element_blank())
GO.plot_calcification_down


ggsave(filename="../../output/WGCNA/GO_analysis/GO.plot_calcification_down.pdf", plot=GO.plot_calcification_down, dpi=300, height=100, units="in", limitsize=FALSE)
```

```{r}
CalcDownMods_GO_P05_reduced <- go_results_BP_reduced %>% 
  mutate(term = fct_reorder(term, over_represented_pvalue, .desc = TRUE)) %>%
  ggplot(aes(x=term, y=over_represented_pvalue) ) +
      geom_segment(aes(x=term ,xend=term, y=0, yend=over_represented_pvalue), color="grey") +
      geom_point(size=1, color="#69b3a2") +
      coord_flip() +
      theme(
        panel.grid.minor.y = element_blank(),
        panel.grid.major.y = element_blank(),
        legend.position="none"
      ) +
  xlab("") +
  ylab("over_represented_pvalue") +
  ggtitle("Biological Process GO Terms Enriched in Modules Downregulated in Calcification") + #add a main title
  theme(plot.title = element_text(face = 'bold',
                                  size = 3,
                                  hjust = 0)) +
  theme_bw() + #Set background color
  theme(panel.border = element_blank(), # Set border
                     panel.grid.major = element_blank(), #Set major gridlines
                     panel.grid.minor = element_blank(), #Set minor gridlines
                     axis.line = element_line(colour = "black"), #Set axes color
                    axis.text.y = element_text(size = 2), #set y axis labels size
        plot.background=element_blank(),#Set the plot background
        legend.position="none")

CalcDownMods_GO_P05_reduced
ggsave("../../output/WGCNA/GO_analysis/CalcDownMods_GO_P05_reduced.pdf", CalcDownMods_GO_P05_reduced, width = 12, height = 36, units = c("in"), dpi=300, limitsize=FALSE)
```

### Calcification - Down - By Module

```{r}
### Generate vector with names of all genes 
ALL.vector <- c(geneInfo$gene_id)
```

```{r}
### Generate length vector for all genes 
LENGTH.vector <- as.integer(geneInfo$Length)
```

#### Turquoise

```{r}
ID.vector <- geneInfo %>%
  filter(moduleColor %in% c("turquoise")) %>%
  pull(gene_id)

length(ID.vector) #2558

##Get a list of GO Terms for the all 9012 genes
GO.terms <- geneInfo %>%
  dplyr::select(gene_id, GOs) %>% rename(GOs = "GO.terms")

dim(GO.terms) #9012    2
```

```{r}
##Format to have one goterm per row with gene ID repeated
split <- strsplit(as.character(GO.terms$GO.terms), ";") 
split2 <- data.frame(v1 = rep.int(GO.terms$gene, sapply(split, length)), v2 = unlist(split)) 
colnames(split2) <- c("gene", "GO.terms")
GO.terms<-split2

##Construct list of genes with 1 for genes in module and 0 for genes not in the module
gene.vector=as.integer(ALL.vector %in% ID.vector) 
names(gene.vector)<-ALL.vector #set names
#weight gene vector by bias for length of gene 
pwf<-nullp(gene.vector, ID.vector, bias.data=LENGTH.vector) 

#run goseq using Wallenius method for all categories of GO terms 
GO.wall<-goseq(pwf, ID.vector, gene2cat=GO.terms, test.cats=c("GO:BP", "GO:MF", "GO:CC"), method="Wallenius", use_genes_without_cat=TRUE)
GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
colnames(GO)[1] <- "GOterm"
```

```{r}
#Filtering for p < 0.05
GO_05_turquoise <- GO %>%
        dplyr::filter(over_represented_pvalue<0.05) %>%
        dplyr::arrange(., ontology, over_represented_pvalue) %>% mutate(Module = "Turquoise")

go_results_BP <- GO_05_turquoise %>%
      filter(ontology=="BP")%>%
      filter(over_represented_pvalue != "NA") %>%
      #filter(numInCat>10)%>%
      arrange(., over_represented_pvalue)

dim(go_results_BP)
```


```{r}
Top10Plot <- go_results_BP %>% 
    top_n(10, wt=-over_represented_pvalue) %>% 
    mutate(hitsPerc=numDEInCat/numInCat) %>% 
    ggplot(aes(x=hitsPerc, 
               y=reorder(term, -hitsPerc), 
               colour=over_represented_pvalue, 
               size=numDEInCat)) +
        geom_point() +
        scale_colour_gradient(low = "red", high = "blue", name = "p value",guide=guide_colorbar(reverse=TRUE)) + 
        expand_limits(x=0) +
        scale_y_discrete(labels = label_wrap(40)) + 
        labs(title = "Top 10 Enriched Biological Process GO Terms, Turquoise Module", x="Proportion of Genes in Module per GO Term", y="GO term", colour="p value", size="Count") +
        theme_bw() +
        theme(axis.text.y = element_text(size = 8), 
              axis.title = element_text(size = 10), 
              plot.title = element_text(hjust = 0.5, size = 10)) 

Top10Plot

ggsave("../../output/WGCNA/GO_analysis/Top10_BP_by_mod/Turquoise.pdf", Top10Plot, width = 5, height = 5, units = c("in"), dpi=300, limitsize=FALSE)
```


Collapsing with simplifyEnrichment package

```{r}
library(simplifyEnrichment)
BP_terms <- go_results_BP$GOterm

Mat <- GO_similarity(BP_terms, ont = "BP", db = "org.Ce.eg.db")
pdf(file = "../../output/WGCNA/GO_analysis/semantic_similarity/calc_down_mods_GOSEM_turquoise.pdf", width = 7, height = 5)
simplifyGO(Mat, word_cloud_grob_param = list(max_width = 50), min_term = 3, exclude_words = c("regulation", "process"), order_by_size = TRUE)
dev.off()
```

```{r}
library(rrvgo)
#Reduce/collapse GO term set with the rrvgo package 
simMatrix <- calculateSimMatrix(go_results_BP$GOterm,
                                orgdb="org.Ce.eg.db", #c. elegans database
                                ont="BP",
                                method="Rel")
```

```{r}
 #calculate similarity 
scores <- setNames(-log(go_results_BP$over_represented_pvalue), go_results_BP$GOterm)
reducedTerms <- reduceSimMatrix(simMatrix,
                                scores,
                                threshold=0.7,
                                orgdb="org.Ce.eg.db")
dim(reducedTerms)
```

```{r}
#keep only the goterms from the reduced list
go_results_BP_reduced <- go_results_BP %>%
  filter(GOterm %in% reducedTerms$go)
 #add in parent terms to list of go terms 
go_results_BP_reduced$ParentTerm <- reducedTerms$parentTerm[match(go_results_BP_reduced$GOterm, reducedTerms$go)]

length(unique(go_results_BP_reduced$ParentTerm))

go_results_BP_reduced <- go_results_BP_reduced %>% group_by(ParentTerm) %>% mutate("N_in_Parent" = n()) %>% ungroup()
```

The reduced list of terms is 291 terms that falls under 46 parent terms.

```{r}
write.csv(go_results_BP_reduced, "../../output/WGCNA/GO_analysis/goseq_turquoise_filtered.csv")
```


What about for p < 0.01?

```{r}
go_results_BP_reduced_01 <- go_results_BP_reduced %>% filter(over_represented_pvalue < 0.01) 
dim(go_results_BP_reduced_01)
length(unique(go_results_BP_reduced_01$ParentTerm))
```

This is only 155 terms that fall into 28 parent terms

```{r}
go_results_BP_reduced_Turquoise <- go_results_BP_reduced_01

top_10_parent <- go_results_BP_reduced_Turquoise %>% dplyr::select(ParentTerm, N_in_Parent) %>% distinct() %>% arrange(-N_in_Parent) %>% head(10)

Top10Plot <- go_results_BP_reduced_Turquoise %>% 
    filter(ParentTerm %in% top_10_parent$ParentTerm) %>%
    mutate(hitsPerc=numDEInCat/numInCat) %>% 
    arrange(-N_in_Parent) %>%
  mutate(ParentTerm = factor(ParentTerm, levels = unique(ParentTerm))) %>%
    ggplot(aes(x=numDEInCat, 
               y=reorder(term, numDEInCat), 
               colour=over_represented_pvalue)) +
        scale_colour_gradient(low = "red", high = "blue", name = "p value",guide=guide_colorbar(reverse=TRUE)) +
        expand_limits(x=1.3) +
        facet_grid(ParentTerm ~ .,scales="free_y",labeller = labeller(ParentTerm = label_wrap_gen(width = 25))) + 
        geom_segment(aes(y=term ,yend=term, x=0, xend=numDEInCat), color="black",linewidth=0.3) +
       geom_point() +
    scale_y_discrete(labels = label_wrap(30)) + 
        labs(title = "Enriched Biological Process GO Terms p < 0.01,\nFacet by Parent Term, Turquoise Module", x="Count", y="GO term", colour="p-value") +
        theme_bw() + 
        theme(axis.text.y = element_text(size = 4), 
              axis.title = element_text(size = 10), 
              plot.title = element_text(hjust = 0.5, size = 10),
              strip.text.y = element_text(size = 6,angle = 0,face = "bold"))

Top10Plot

ggsave("../../output/WGCNA/GO_analysis/Top10_BP_by_mod/Turquoise_allterms_top10parent_filtp01.pdf", Top10Plot, width = 5, height = 7, units = c("in"), dpi=300, limitsize=FALSE)

top_5_parent <- go_results_BP_reduced_Turquoise %>% dplyr::select(ParentTerm, N_in_Parent) %>% distinct() %>% arrange(-N_in_Parent) %>% head(5)

Top5Plot <- go_results_BP_reduced_Turquoise %>% 
    filter(ParentTerm %in% top_5_parent$ParentTerm) %>%
    filter(over_represented_pvalue <0.01) %>%
    mutate(hitsPerc=numDEInCat/numInCat) %>% 
    arrange(-N_in_Parent) %>%
  mutate(ParentTerm = factor(ParentTerm, levels = unique(ParentTerm))) %>%
    
    ggplot(aes(x=numDEInCat, 
               y=reorder(term, numDEInCat), 
               colour=over_represented_pvalue)) +
        scale_colour_gradient(low = "red", high = "blue", name = "p value",guide=guide_colorbar(reverse=TRUE)) +
        expand_limits(x=1.3) +
        facet_grid(ParentTerm ~ .,scales="free_y", space = "free",labeller = labeller(ParentTerm = label_wrap_gen(width = 25))) + 
        geom_segment(aes(y=term ,yend=term, x=0, xend=numDEInCat), color="black",linewidth=0.3) +
       geom_point() +
    scale_y_discrete(labels = label_wrap(30)) + 
        labs(title = "Enriched Biological Process GO Terms p < 0.01,\nFacet by Parent Term (Top 5), Turquoise Module", x="Count", y="GO term", colour="p-value") +
        theme_bw() + 
        theme(axis.text.y = element_text(size = 4), 
              axis.title = element_text(size = 10), 
              plot.title = element_text(hjust = 0.5, size = 10),
              strip.text.y = element_text(size = 6,angle = 0,face = "bold"))

Top5Plot

ggsave("../../output/WGCNA/GO_analysis/Top10_BP_by_mod/Turquoise_allterms_top5parent_filtp01.pdf", Top5Plot, width = 5, height = 5, units = c("in"), dpi=300, limitsize=FALSE)
```


#### Magenta

```{r}
ID.vector <- geneInfo %>%
  filter(moduleColor %in% c("magenta")) %>%
  pull(gene_id)

length(ID.vector) #219

##Get a list of GO Terms for the all 9012 genes
GO.terms <- geneInfo %>%
  dplyr::select(gene_id, GOs) %>% rename(GOs = "GO.terms")

dim(GO.terms) #9012    2
```

```{r}
##Format to have one goterm per row with gene ID repeated
split <- strsplit(as.character(GO.terms$GO.terms), ";") 
split2 <- data.frame(v1 = rep.int(GO.terms$gene, sapply(split, length)), v2 = unlist(split)) 
colnames(split2) <- c("gene", "GO.terms")
GO.terms<-split2

##Construct list of genes with 1 for genes in module and 0 for genes not in the module
gene.vector=as.integer(ALL.vector %in% ID.vector) 
names(gene.vector)<-ALL.vector#set names
#weight gene vector by bias for length of gene 
pwf<-nullp(gene.vector, ID.vector, bias.data=LENGTH.vector) 

#run goseq using Wallenius method for all categories of GO terms 
GO.wall<-goseq(pwf, ID.vector, gene2cat=GO.terms, test.cats=c("GO:BP", "GO:MF", "GO:CC"), method="Wallenius", use_genes_without_cat=TRUE)
GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
colnames(GO)[1] <- "GOterm"
```

```{r}
#Filtering for p < 0.05
GO_05_magenta <- GO %>%
        dplyr::filter(over_represented_pvalue<0.05) %>%
        dplyr::arrange(., ontology, over_represented_pvalue) %>% mutate(Module = "Magenta")

go_results_BP <- GO_05_magenta %>%
      filter(ontology=="BP")%>%
      filter(over_represented_pvalue != "NA") %>%
      #filter(numInCat>10)%>%
      arrange(., over_represented_pvalue)

dim(go_results_BP)
```

```{r}
#Write file of results 
write.csv(GO_05_magenta, file = "../../output/WGCNA/GO_analysis/goseq_pattern_calcification_magenta.csv")
```


```{r}
Top10Plot <- go_results_BP %>% 
    top_n(10, wt=-over_represented_pvalue) %>% 
    mutate(hitsPerc=numDEInCat/numInCat) %>% 
    ggplot(aes(x=hitsPerc, 
               y=reorder(term, -hitsPerc), 
               colour=over_represented_pvalue, 
               size=numDEInCat)) +
        geom_point() +
        scale_colour_gradient(low = "red", high = "blue", name = "p value",guide=guide_colorbar(reverse=TRUE)) + 
        expand_limits(x=0) +
        scale_y_discrete(labels = label_wrap(40)) + 
        labs(title = "Top 10 Enriched Biological Process GO Terms, Magenta Module", x="Proportion of Genes in Module per GO Term", y="GO term", colour="p value", size="Count") +
        theme_bw() + 
        theme(axis.text.y = element_text(size = 8), 
              axis.title = element_text(size = 10), 
              plot.title = element_text(hjust = 0.5, size = 10)) 

Top10Plot

ggsave("../../output/WGCNA/GO_analysis/Top10_BP_by_mod/Magenta.pdf", Top10Plot, width = 5, height = 5, units = c("in"), dpi=300, limitsize=FALSE)
```

Collapsing with simplifyEnrichment package

```{r}
library(simplifyEnrichment)
BP_terms <- go_results_BP$GOterm

Mat <- GO_similarity(BP_terms, ont = "BP", db = "org.Ce.eg.db")
pdf(file = "../../output/WGCNA/GO_analysis/semantic_similarity/calc_down_mods_GOSEM_magenta.pdf", width = 7, height = 5)
result <-  simplifyGO(Mat, word_cloud_grob_param = list(max_width = 50),  min_term = 3, exclude_words = c("regulation", "process"), order_by_size = TRUE)
result
dev.off()

# Following filtering of small clusters is based on the simplifyGO/ht_clusters source code for merging the small clusters together

# Filter clusters smaller than 3
cluster_sizes <- table(result$cluster) # table function makes a table with the size of each cluster (freqeuncy) 
large_clusters <- as.numeric(names(cluster_sizes)[cluster_sizes >= 3]) # Identify clusters with frequenct/size less than 3
filtered_indices <- result$cluster %in% large_clusters # Filter out the indices belonging to small clusters
filtered_Mat <- Mat[filtered_indices, filtered_indices] # Create a new matrix with only the indicies from larger clusters


pdf(file = "../../output/WGCNA/GO_analysis/semantic_similarity/calc_down_mods_GOSEM_magenta_filtered.pdf", width = 7, height = 5)
simplifyGO(filtered_Mat, word_cloud_grob_param = list(max_width = 50),  min_term = 1, exclude_words = c("regulation", "process"), order_by_size = TRUE)
dev.off()
```


```{r}
library(rrvgo)
#Reduce/collapse GO term set with the rrvgo package 
simMatrix <- calculateSimMatrix(go_results_BP$GOterm,
                                orgdb="org.Ce.eg.db", #c. elegans database
                                ont="BP",
                                method="Rel")
```

```{r}
 #calculate similarity 
scores <- setNames(-log(go_results_BP$over_represented_pvalue), go_results_BP$GOterm)
reducedTerms <- reduceSimMatrix(simMatrix,
                                scores,
                                threshold=0.7,
                                orgdb="org.Ce.eg.db")
dim(reducedTerms)
```

```{r}
#keep only the goterms from the reduced list
go_results_BP_reduced <- go_results_BP %>%
  filter(GOterm %in% reducedTerms$go)
 #add in parent terms to list of go terms 
go_results_BP_reduced$ParentTerm <- reducedTerms$parentTerm[match(go_results_BP_reduced$GOterm, reducedTerms$go)]

length(unique(go_results_BP_reduced$ParentTerm))

go_results_BP_reduced <- go_results_BP_reduced %>% group_by(ParentTerm) %>% mutate("N_in_Parent" = n()) %>% ungroup()
```

The reduced list of terms is 79 terms that falls under 23 parent terms.

```{r}
write.csv(go_results_BP_reduced, "../../output/WGCNA/GO_analysis/goseq_magenta_filtered.csv")
```


What about for p < 0.01?

```{r}
go_results_BP_reduced_01 <- go_results_BP_reduced %>% filter(over_represented_pvalue < 0.01) 
dim(go_results_BP_reduced_01)
length(unique(go_results_BP_reduced_01$ParentTerm))
```

This is only 14 terms that fall into 6 parent terms

```{r}
go_results_BP_reduced_Magenta <- go_results_BP_reduced_01

top_10_parent <- go_results_BP_reduced_Magenta %>% dplyr::select(ParentTerm, N_in_Parent) %>% distinct() %>% arrange(-N_in_Parent) %>% head(10)

Top10Plot <- go_results_BP_reduced_Magenta %>% 
    filter(ParentTerm %in% top_10_parent$ParentTerm) %>%
    mutate(hitsPerc=numDEInCat/numInCat) %>% 
    arrange(-N_in_Parent) %>%
  mutate(ParentTerm = factor(ParentTerm, levels = unique(ParentTerm))) %>%
    ggplot(aes(x=numDEInCat, 
               y=reorder(term, numDEInCat), 
               colour=over_represented_pvalue)) +
        scale_colour_gradient(low = "red", high = "blue", name = "p value",guide=guide_colorbar(reverse=TRUE)) +
        expand_limits(x=1.3) +
        facet_grid(ParentTerm ~ .,scales="free_y",labeller = labeller(ParentTerm = label_wrap_gen(width = 25))) + 
        geom_segment(aes(y=term ,yend=term, x=0, xend=numDEInCat), color="black",linewidth=0.3) +
       geom_point() +
    scale_y_discrete(labels = label_wrap(30)) + 
        labs(title = "Enriched Biological Process GO Terms p < 0.01,\nFacet by Parent Term, Magenta Module", x="Count", y="GO term", colour="p-value") +
        theme_bw() + 
        theme(axis.text.y = element_text(size = 4), 
              axis.title = element_text(size = 10), 
              plot.title = element_text(hjust = 0.5, size = 10),
              strip.text.y = element_text(size = 6,angle = 0,face = "bold"))

Top10Plot

ggsave("../../output/WGCNA/GO_analysis/Top10_BP_by_mod/Magenta_allterms_top10parent_filtp01.pdf", Top10Plot, width = 5, height = 7, units = c("in"), dpi=300, limitsize=FALSE)

top_5_parent <- go_results_BP_reduced_Magenta %>% dplyr::select(ParentTerm, N_in_Parent) %>% distinct() %>% arrange(-N_in_Parent) %>% head(10)

Top5Plot <- go_results_BP_reduced_Magenta %>% 
    filter(ParentTerm %in% top_5_parent$ParentTerm) %>%
    filter(over_represented_pvalue <0.01) %>%
    mutate(hitsPerc=numDEInCat/numInCat) %>% 
    arrange(-N_in_Parent) %>%
  mutate(ParentTerm = factor(ParentTerm, levels = unique(ParentTerm))) %>%
    
    ggplot(aes(x=numDEInCat, 
               y=reorder(term, numDEInCat), 
               colour=over_represented_pvalue)) +
        scale_colour_gradient(low = "red", high = "blue", name = "p value",guide=guide_colorbar(reverse=TRUE)) +
        expand_limits(x=1.3) +
        facet_grid(ParentTerm ~ .,scales="free_y", space = "free",labeller = labeller(ParentTerm = label_wrap_gen(width = 25))) + 
        geom_segment(aes(y=term ,yend=term, x=0, xend=numDEInCat), color="black",linewidth=0.3) +
       geom_point() +
    scale_y_discrete(labels = label_wrap(30)) + 
        labs(title = "Enriched Biological Process GO Terms p < 0.01,\nFacet by Parent Term (Top 5), Magenta Module", x="Count", y="GO term", colour="p-value") +
        theme_bw() + 
        theme(axis.text.y = element_text(size = 4), 
              axis.title = element_text(size = 10), 
              plot.title = element_text(hjust = 0.5, size = 10),
              strip.text.y = element_text(size = 6,angle = 0,face = "bold"))

Top5Plot

ggsave("../../output/WGCNA/GO_analysis/Top10_BP_by_mod/Magenta_allterms_top5parent_filtp01.pdf", Top5Plot, width = 5, height = 5, units = c("in"), dpi=300, limitsize=FALSE)
```


#### Lightcyan

```{r}
ID.vector <- geneInfo %>%
  filter(moduleColor %in% c("lightcyan")) %>%
  pull(gene_id)

length(ID.vector) #65

##Get a list of GO Terms for the all 9012 genes
GO.terms <- geneInfo %>%
  dplyr::select(gene_id, GOs) %>% rename(GOs = "GO.terms")

dim(GO.terms) #9012    2
```

```{r}
##Format to have one goterm per row with gene ID repeated
split <- strsplit(as.character(GO.terms$GO.terms), ";") 
split2 <- data.frame(v1 = rep.int(GO.terms$gene, sapply(split, length)), v2 = unlist(split)) 
colnames(split2) <- c("gene", "GO.terms")
GO.terms<-split2

##Construct list of genes with 1 for genes in module and 0 for genes not in the module
gene.vector=as.integer(ALL.vector %in% ID.vector) 
names(gene.vector)<-ALL.vector#set names
#weight gene vector by bias for length of gene 
pwf<-nullp(gene.vector, ID.vector, bias.data=LENGTH.vector) 

#run goseq using Wallenius method for all categories of GO terms 
GO.wall<-goseq(pwf, ID.vector, gene2cat=GO.terms, test.cats=c("GO:BP", "GO:MF", "GO:CC"), method="Wallenius", use_genes_without_cat=TRUE)
GO <- GO.wall[order(GO.wall$over_represented_pvalue),]
colnames(GO)[1] <- "GOterm"
```

```{r}
#Filtering for p < 0.05
GO_05_lightcyan <- GO %>%
        dplyr::filter(over_represented_pvalue<0.05) %>%
        dplyr::arrange(., ontology, over_represented_pvalue) %>% mutate(Module = "Lightcyan")

go_results_BP <- GO_05_lightcyan %>%
      filter(ontology=="BP")%>%
      filter(over_represented_pvalue != "NA") %>%
      #filter(numInCat>10)%>%
      arrange(., over_represented_pvalue)

dim(go_results_BP)
```

```{r}
Top10Plot <- go_results_BP %>% 
    top_n(10, wt=-over_represented_pvalue) %>% 
    mutate(hitsPerc=numDEInCat/numInCat) %>% 
    ggplot(aes(x=hitsPerc, 
               y=reorder(term, -hitsPerc), 
               colour=over_represented_pvalue, 
               size=numDEInCat)) +
        geom_point() +
        scale_colour_gradient(low = "red", high = "blue", name = "p value",guide=guide_colorbar(reverse=TRUE)) + 
        expand_limits(x=0) +
        scale_y_discrete(labels = label_wrap(40)) + 
        labs(title = "Top 10 Enriched Biological Process GO Terms, Lightcyan Module", x="Proportion of Genes in Module per GO Term", y="GO term", colour="p value", size="Count") +
        theme_bw() + 
        theme(axis.text.y = element_text(size = 8), 
              axis.title = element_text(size = 10), 
              plot.title = element_text(hjust = 0.5, size = 10)) 

Top10Plot

ggsave("../../output/WGCNA/GO_analysis/Top10_BP_by_mod/Lightcyan.pdf", Top10Plot, width = 5, height = 5, units = c("in"), dpi=300, limitsize=FALSE)
```

Collapsing with simplifyEnrichment package

```{r}
library(simplifyEnrichment)
BP_terms <- go_results_BP$GOterm

Mat <- GO_similarity(BP_terms, ont = "BP", db = "org.Ce.eg.db")
pdf(file = "../../output/WGCNA/GO_analysis/semantic_similarity/calc_down_mods_GOSEM_lightcyan.pdf", width = 7, height = 5)
simplifyGO(Mat, word_cloud_grob_param = list(max_width = 50), min_term = 3, exclude_words = c("regulation", "process"), order_by_size = TRUE)
dev.off()
```


### Combining all the individual module lists for Calcification - Down

```{r}
GO_05_DOWN <- rbind(GO_05_turquoise,GO_05_magenta,GO_05_lightcyan)

length(GO_05_DOWN$GOterm) #1072 enriched GO terms
length(unique(GO_05_DOWN$GOterm)) #1070 enriched GO terms

# Collapse GO terms that are duplicated accross modules - sort by GO term and p-value, then keep the row with the lowest p-value, collapsing the module names so you know which modules were represented

# Identify and collapse duplicates
GO_05_DOWN <- GO_05_DOWN %>%
  arrange(GOterm, over_represented_pvalue) %>%
  group_by(GOterm) %>% mutate(Module = paste(unique(Module), collapse = ";")) %>%
  group_by(GOterm) %>%
  slice_min(order_by = over_represented_pvalue, n = 1) %>%
  ungroup()

#Write file of results 
write.csv(GO_05_DOWN, file = "../../output/WGCNA/GO_analysis/goseq_pattern_calcification_down_bymod.csv")
```

## Merge the frequency of GOterms for both up- and down-regulation of calcification

### Reduce the lists together
```{r}
# GO_05_up <- read.csv("../../output/WGCNA/GO_analysis/goseq_pattern_calcification.csv") %>% dplyr::select(-"X")
# GO_05_down <- read.csv("../../output/WGCNA/GO_analysis/goseq_pattern_calcification_down.csv") %>% dplyr::select(-"X")

GO_05_up <- read.csv("../../output/WGCNA/GO_analysis/goseq_pattern_calcification_up_bymod.csv") %>% dplyr::select(-"X")
GO_05_down <- read.csv("../../output/WGCNA/GO_analysis/goseq_pattern_calcification_down_bymod.csv") %>% dplyr::select(-"X")
```

```{r}
go_results_BP_up <- GO_05_up %>%
      filter(ontology=="BP")%>%
      filter(over_represented_pvalue != "NA") %>%
      #filter(numInCat>10)%>%
      arrange(., over_represented_pvalue)
dim(go_results_BP_up)

go_results_BP_down <- GO_05_down %>%
      filter(ontology=="BP")%>%
      filter(over_represented_pvalue != "NA") %>%
      #filter(numInCat>10)%>%
      arrange(., over_represented_pvalue)
dim(go_results_BP_down)

all_enriched_terms_up_down <- c(go_results_BP_up$GOterm,go_results_BP_down$GOterm)
length(all_enriched_terms_up_down)
length(unique(all_enriched_terms_up_down))
```

By reducing the lists together, we can ensure that each GO term is given the same parent term whether the term was in the Up or Down modules.

```{r}
library(rrvgo)
#Reduce/collapse GO term set with the rrvgo package 
simMatrix <- calculateSimMatrix(unique(all_enriched_terms_up_down),
                                orgdb="org.Ce.eg.db", #c. elegans database
                                ont="BP",
                                method="Rel")
```

```{r}
#calculate similarity 
reducedTerms <- reduceSimMatrix(simMatrix,
                                scores = "size",
                                threshold=0.7,
                                orgdb="org.Ce.eg.db")
dim(reducedTerms)
```

```{r}
#keep only the goterms from the reduced list
go_results_BP_up_reduced <- go_results_BP_up %>%
  filter(GOterm %in% reducedTerms$go)
 #add in parent terms to list of go terms 
go_results_BP_up_reduced$ParentTerm <- reducedTerms$parentTerm[match(go_results_BP_up_reduced$GOterm, reducedTerms$go)]

length(unique(go_results_BP_up_reduced$ParentTerm))

#keep only the goterms from the reduced list
go_results_BP_down_reduced <- go_results_BP_down %>%
  filter(GOterm %in% reducedTerms$go)
 #add in parent terms to list of go terms 
go_results_BP_down_reduced$ParentTerm <- reducedTerms$parentTerm[match(go_results_BP_down_reduced$GOterm, reducedTerms$go)]

length(unique(go_results_BP_down_reduced$ParentTerm))
```

```{r}
#write.csv(go_results_BP_up_reduced, "../../output/WGCNA/GO_analysis/goseq_pattern_calcification_filtered.csv")
#write.csv(go_results_BP_down_reduced, "../../output/WGCNA/GO_analysis/goseq_pattern_calcification_down_filtered.csv")
```

The following list is a list of BP-ontology ("Biological Process") GO terms that were significantly enriched (P < 0.05) within the modules that were upregulated for calcification. The list has been further reduced by using the package rrvgo.

```{r}
#cal_up_terms <- read.csv("../../output/WGCNA/GO_analysis/goseq_pattern_calcification_filtered.csv")
cal_up_terms <- go_results_BP_up_reduced
cal_up_terms <- cal_up_terms %>% mutate(Factor = "Up") 
head(cal_up_terms)
```

The following list is a list of BP-ontology ("Biological Process") GO terms that were significantly enriched (P < 0.05) within the modules that were downregulated for calcification. The list has been further reduced by using the package rrvgo.

```{r}
#cal_down_terms <- read.csv("../../output/WGCNA/GO_analysis/goseq_pattern_calcification_down_filtered.csv")
cal_down_terms <- go_results_BP_down_reduced
cal_down_terms <- cal_down_terms %>% mutate(Factor = "Down")
head(cal_down_terms)
```

### Merge up and down-regulation of calcification GOterms

The following list is a list of BP-ontology ("Biological Process") GO terms that were significantly enriched (P < 0.05) within the modules that were upregulated *or* downregulated for calcification. The list has been further reduced by using the package rrvgo.

```{r}
all_terms <- rbind(cal_down_terms,cal_up_terms)

all_terms <-  all_terms[,c("Factor","GOterm","over_represented_pvalue","under_represented_pvalue","numDEInCat","numInCat","term","ontology","ParentTerm","Module")]

all_terms$GOterm<-as.factor(all_terms$GOterm)

dim(all_terms) #2140 reduced terms 

length(unique(all_terms$GOterm)) #this represents 2004 unique terms between the two lists of reduced terms
length(unique(all_terms$ParentTerm)) #this represents 153 unique terms between the two lists of reduced terms
```

### Unique terms

This is collapsing the list from 2236 rows to 2004, representing the 2004 unique GO terms in the list above. It collapses the information in the columns "ParentTerm" and "Factor" so that for each GO term we get the parent terms associated and if this was enriched in modules upregulated for calcification, downregulated for calcification, or both.

```{r}
goterms_shared <- all_terms %>%
  group_by(GOterm) %>%
  dplyr::summarise(
    ParentTerm = paste(unique(ParentTerm), collapse = ", "),
    Factor = paste(unique(Factor), collapse = ", "),
    term = paste(unique(term), collapse = ", ")
  )
dim(goterms_shared)

goterms_shared_full <- goterms_shared %>%
  left_join(dplyr::select(all_terms,GOterm, Factor, over_represented_pvalue), by = "GOterm") %>% #select 3 columns GOterm, Factor, over_represented_pvalue from all_terms and left join by GOterm, turns this dataframe from 2004 rows back to the 2236 in all_terms
  mutate(pvalue_Up = ifelse(Factor.y == "Up", over_represented_pvalue, NA)) %>% #add a column that is the p-value for the Up factor
  mutate(pvalue_Down = ifelse(Factor.y == "Down", over_represented_pvalue, NA)) %>% #add a column that is the p-value for the Down factor
  dplyr::select(-c("over_represented_pvalue", "Factor.y")) %>% #remove the unique columns so that we can collapse the dataframe
  group_by(GOterm, ParentTerm, Factor.x, term) %>% #group by the repeated columns for the non-unique GO terms
  dplyr::summarize(pvalue_Up = na.omit(pvalue_Up)[1], #carry over the p-value for the term in the up direction, by taking the first non-NA value.
            pvalue_Down = na.omit(pvalue_Down)[1]) %>% #carry over the p-value for the term in the down direction, by taking the first non-NA value.
  rename(Factor.x = "Factor") #rename column 
```

```{r}
write.csv(goterms_shared_full, "../../output/WGCNA/GO_analysis/Merged_GOterms_factor_ParentTerm.csv")
```

### plots

```{r}
goterms_shared_full <- goterms_shared_full %>% group_by(ParentTerm) %>% mutate("N_in_Parent" = n()) %>% ungroup()

goterms_up <- goterms_shared_full %>% dplyr::filter(Factor=="Up")
goterms_down <- goterms_shared_full %>% dplyr::filter(Factor=="Down")
goterms_shared_only <- goterms_shared_full %>% dplyr::filter(Factor=="Down, Up")

nrow(goterms_up)
nrow(goterms_down)
nrow(goterms_shared_only)

nrow(goterms_up) + nrow(goterms_down) + nrow(goterms_shared_only) == nrow(goterms_shared_full)

#how many parent terms in each
length(unique(goterms_up$ParentTerm))
length(unique(goterms_down$ParentTerm))
length(unique(goterms_shared_only$ParentTerm))
```
  
```{r}
freq_fig_up_pval <- ggplot(goterms_shared_only, aes(x = reorder(ParentTerm, -N_in_Parent), y = pvalue_Up)) +
  geom_segment(aes(x = ParentTerm, xend = ParentTerm, y = 0, yend = pvalue_Up,),linewidth=.25) +
  geom_point(size=1.5, alpha=1, colour = "red") +
  geom_hline(yintercept = 0, linetype = "solid", color = "grey", size = 0.5, show.legend = TRUE) +
  coord_flip() +
  theme_classic() +
  ylim(0, 0.05) + 
  theme(axis.text.x=element_text(vjust=0.5, hjust=0.95,size=5),
        plot.title = element_text(margin = margin(t = 10, b = 10), hjust=0.5),
        panel.background= element_rect(fill=NA, color='black'),
        axis.text.y = element_text(vjust=0.5, size=5), 
        axis.title.x = element_blank(),
        axis.title.y = element_blank()) +
  guides(color = FALSE)

freq_fig_down_pval <- ggplot(goterms_shared_only, aes(x = reorder(ParentTerm, -N_in_Parent), y = pvalue_Down)) +
  geom_segment(aes(x = ParentTerm, xend = ParentTerm, y = 0, yend = pvalue_Down),linewidth=.25) +
  geom_point(size=1.5, alpha=1, colour = "blue") +
  geom_hline(yintercept = 0, linetype = "solid", color = "grey", size = 0.5, show.legend = TRUE) +
  coord_flip() +
  theme_classic() +
  ylim(0, 0.05) + 
  theme(axis.text.x=element_text(vjust=0.5, hjust=0.95,size=5),
        plot.title = element_text(margin = margin(t = 10, b = 10), hjust=0.5),
        panel.background= element_rect(fill=NA, color='black'),
        axis.text.y = element_text(vjust=0.5, size=5), 
        axis.title.x = element_blank(),
        axis.title.y = element_blank()) 

compare_figs<-cowplot::plot_grid(freq_fig_up_pval, freq_fig_down_pval, nrow=1, align="v")
compare_figs

ggsave(filename="../../output/WGCNA/GO_analysis/Up_Down_Shared_ParentTerms.pdf", plot=compare_figs, dpi=300, height=3, units="in", limitsize=FALSE)
```

```{r}
freq_fig_up_pval <- ggplot(goterms_up, aes(x = reorder(ParentTerm, -N_in_Parent), y = pvalue_Up)) +
  geom_segment(aes(x = ParentTerm, xend = ParentTerm, y = 0, yend = pvalue_Up,),linewidth=.25) +
  geom_point(size=1.5, alpha=1, colour = "red") +
  geom_hline(yintercept = 0, linetype = "solid", color = "grey", size = 0.5, show.legend = TRUE) +
  coord_flip() +
  theme_classic() +
  ylim(0, 0.05) + 
  theme(axis.text.x=element_text(vjust=0.5, hjust=0.95,size=5),
        plot.title = element_text(margin = margin(t = 10, b = 10), hjust=0.5),
        panel.background= element_rect(fill=NA, color='black'),
        axis.text.y = element_text(vjust=0.5, size=5), 
        axis.title.x = element_blank(),
        axis.title.y = element_blank()) +
  guides(color = FALSE)

freq_fig_down_pval <- ggplot(goterms_down, aes(x = reorder(ParentTerm, -N_in_Parent), y = pvalue_Down)) +
  geom_segment(aes(x = ParentTerm, xend = ParentTerm, y = 0, yend = pvalue_Down),linewidth=.25) +
  geom_point(size=1.5, alpha=1, colour = "blue") +
  geom_hline(yintercept = 0, linetype = "solid", color = "grey", size = 0.5, show.legend = TRUE) +
  coord_flip() +
  theme_classic() +
  ylim(0, 0.05) + 
  theme(axis.text.x=element_text(vjust=0.5, hjust=0.95,size=5),
        plot.title = element_text(margin = margin(t = 10, b = 10), hjust=0.5),
        panel.background= element_rect(fill=NA, color='black'),
        axis.text.y = element_text(vjust=0.5, size=5), 
        axis.title.x = element_blank(),
        axis.title.y = element_blank())

compare_figs<-cowplot::plot_grid(freq_fig_up_pval, freq_fig_down_pval, nrow=1, align="v")
compare_figs

ggsave(filename="../../output/WGCNA/GO_analysis/Up_Down_Unique_ParentTerms.pdf", plot=compare_figs, dpi=300, height=10, units="in", limitsize=FALSE)
```

### GOSlim

I downloaded the GOSlim terms list from the following website on 4/5/2024: http://www.informatics.jax.org/gotools/data/input/map2MGIslim.txt

I did so by going to [this link](http://www.informatics.jax.org/gotools/MGI_GO_Slim.html) and right clicking "here" at "Tab-delimited file mapping GO_ids to MGI GO_Slim category available for download _here_" and saying "Save link as..." , and I saved it as "map2MGIslim.txt". I then converted this to CSV and uploaded to this github repository [here](https://github.com/imkristenbrown/Heron-Pdam-gene-expression/blob/master/BioInf/data/map2MGIslim.csv).


Add slim terms to dataset 
```{r}
GOslim <- read.csv("../../data/map2MGIslim.csv") %>% dplyr::select(-c(term,aspect))
```

```{r}
goterms_shared_full_slim <- goterms_shared_full %>%
  inner_join(GOslim, by = c("GOterm" = "GO_id"))

goterms_up_slim <- goterms_shared_full_slim %>% dplyr::filter(Factor=="Up" | Factor=="Down, Up")

goterms_up_full_slim_plot <- goterms_up_slim %>%
      ggplot(aes(x=pvalue_Up, 
               y=term)) +
        geom_point(size = 0.5) +
        expand_limits(x=0) +
        facet_grid(GOSlim_bin ~ ., scales = "free", space='free', labeller = label_wrap_gen(width = 55, multi_line = TRUE)) +
      labs(x="Over-represented p-value", y="") +
      theme_bw() + 
      theme(
        strip.text.y = element_text(size = 5, colour = "black", angle = 0, face = "bold"),
        axis.text.y = element_text(size = 1, colour = "black"),
        axis.title.x = element_text(size = 5, colour = "black"),
      ); goterms_up_full_slim_plot

ggsave("../../output/WGCNA/GO_analysis/Calc_Up_GOSlim.pdf", goterms_up_full_slim_plot, width = 5, height = 35)

goterms_up_full_slim_plot <- goterms_up_slim %>%
      ggplot(aes(x=log10(pvalue_Up), 
               y=term)) +
        geom_point(size = 0.5) +
        expand_limits(x=0) +
        facet_grid(GOSlim_bin ~ ., scales = "free", space='free', labeller = label_wrap_gen(width = 55, multi_line = TRUE)) +
      labs(x="Log 10(Over-represented p-value)", y="") +
      theme_bw() + 
      theme(
        strip.text.y = element_text(size = 5, colour = "black", angle = 0, face = "bold"),
        axis.text.y = element_text(size = 1, colour = "black"),
        axis.title.x = element_text(size = 5, colour = "black"),
      ); goterms_up_full_slim_plot

ggsave("../../output/WGCNA/GO_analysis/Calc_Up_GOSlim_Log.pdf", goterms_up_full_slim_plot, width = 5, height = 35)
```

```{r}
goterms_down_slim <- goterms_shared_full_slim %>% dplyr::filter(Factor=="Down" | Factor=="Down, Up")

goterms_down_full_slim_plot <- goterms_down_slim %>%
      ggplot(aes(x=pvalue_Down, 
               y=term)) +
        geom_point(size = 0.5) +
        expand_limits(x=0) +
        facet_grid(GOSlim_bin ~ ., scales = "free", space='free', labeller = label_wrap_gen(width = 55, multi_line = TRUE)) +
      labs(x="Over-represented p-value", y="") +
      theme_bw() + 
      theme(
        strip.text.y = element_text(size = 5, colour = "black", angle = 0, face = "bold"),
        axis.text = element_text(size = 1, colour = "black"),
        axis.title.x = element_text(size = 5, colour = "black"),
        legend.text=element_text(size = 5),
        legend.title = element_text(size = 5)
      ); goterms_down_full_slim_plot

ggsave("../../output/WGCNA/GO_analysis/Calc_Down_GOSlim.pdf", goterms_down_full_slim_plot, width = 5, height = 35)

goterms_down_full_slim_plot <- goterms_down_slim %>%
      ggplot(aes(x=log10(pvalue_Down), 
               y=term)) +
        geom_point(size = 0.5) +
        expand_limits(x=0) +
        facet_grid(GOSlim_bin ~ ., scales = "free", space='free', labeller = label_wrap_gen(width = 55, multi_line = TRUE)) +
      labs(x="Log 10(Over-represented p-value)", y="") +
      theme_bw() + 
      theme(
        strip.text.y = element_text(size = 5, colour = "black", angle = 0, face = "bold"),
        axis.text = element_text(size = 1, colour = "black"),
        axis.title.x = element_text(size = 5, colour = "black"),
        legend.text=element_text(size = 5),
        legend.title = element_text(size = 5)
      ); goterms_down_full_slim_plot

ggsave("../../output/WGCNA/GO_analysis/Calc_Down_GOSlim_Log.pdf", goterms_down_full_slim_plot, width = 5, height = 35)
```


### For each parent term in this list, how many are "Up" by calcification, "Down", or both

SharedGOterms = # of GO terms within the parent term that are in the list for "Down", "Up", or "Down, Up"
```{r}
result_parent_unique <- goterms_shared %>%
  group_by(ParentTerm,Factor) %>%
  dplyr::summarise(SharedGOterms = n_distinct(GOterm)) %>%
  arrange(-SharedGOterms)
```

```{r}
parent_filtered_up <- result_parent_unique %>%
  dplyr::filter(Factor=="Up" | Factor=="Down, Up")
 #dplyr::filter(SharedGOterms>=5) 
```

```{r}
parent_filtered_down <- result_parent_unique %>%
  dplyr::filter(Factor=="Down" | Factor=="Down, Up")
 #dplyr::filter(SharedGOterms>=5) 
```

### Count number of GOterms by ParentTerm for the upregulation of calcification
```{r}
result_up <- cal_up_terms %>%
  dplyr::group_by(ParentTerm) %>%
  dplyr::summarize(Number.of.terms = n_distinct(term))%>%
  mutate(Calcification.direction = "Up")

head(result_up)
dim(result_up)
```

```{r}
merged_up <- parent_filtered_up %>%
  tidyr::separate_rows(ParentTerm, sep = ", ") %>%
  dplyr::group_by(ParentTerm) %>%
  dplyr::summarize(Sum_of_SharedGOterms = sum(SharedGOterms, na.rm = TRUE)) %>%
  left_join(result_up, by = "ParentTerm") #join with result from above to get the Number.of.Terms column

merged_up_clean <- na.omit(merged_up)
head(merged_up_clean)

dim(merged_up_clean)
```

### Count number of GOterms by ParentTerm for the downregulation of calcification

```{r}
result_down <- cal_down_terms %>%
  dplyr::group_by(ParentTerm) %>%
  dplyr::summarize(Number.of.terms = n_distinct(term))%>%
  mutate(Calcification.direction = "Down")

dim(result_down)
head(result_down)
```

```{r}
merged_down <- parent_filtered_down %>%
  tidyr::separate_rows(ParentTerm, sep = ", ") %>%
  dplyr::group_by(ParentTerm) %>%
  dplyr::summarize(Sum_of_SharedGOterms = sum(SharedGOterms, na.rm = TRUE)) %>%
  left_join(result_down, by = "ParentTerm") 

merged_down_clean<- na.omit(merged_down)
head(merged_down_clean)
```

### Frequency >10 up
```{r}
cal_freq_terms_filtered_up <- merged_up_clean %>%
  filter(Number.of.terms>=10) %>%
  filter(Calcification.direction=="Up")
dim(cal_freq_terms_filtered_up)
```

#### Figure
```{r}
#counts$Direction.of.flat.origin<- factor(counts$Direction.of.flat.origin, levels =c("up","no pattern","down"))
#counts$Module<- factor(counts$Module, levels=c("Blue","Brown","Greenyellow","Cyan","Pink","Magenta","Lightcyan","Midnight blue","Purple","Turquiose","Red","Black"))
freq_fig_up<-ggplot(cal_freq_terms_filtered_up, aes(y=Number.of.terms,x=reorder(ParentTerm, Number.of.terms), group=1))+
  #facet_wrap(~Calcification.direction, nrow = 1)+
  geom_point(size=5, alpha=1, pch=21,color="black")+
  geom_segment(aes(x=ParentTerm, xend=ParentTerm, y=0, yend=Number.of.terms)) +
  geom_hline(yintercept = 0, linetype="solid", color = 'black', size=0.5, show.legend = TRUE)+
  coord_flip()+
  scale_y_continuous(expression(GO~term~counts),limits=c(0,60))+
  #scale_color_manual("Direction.of.flat.origin",values= c("up"="#b2182b","no pattern"="grey","down" ="#67a9cf"))+
  #scale_fill_manual("Direction.of.flat.origin",values= c("up"="#b2182b","no pattern"="grey","down" ="#67a9cf"))+
  scale_fill_gradientn(colours=c("white","#fddbc7","#f4a582","#d6604d","#b2182b"), na.value = "grey98",limits = c(0, 100))+ 
 #scale_color_gradientn(colours=c("#b2182b","#fddbc7","white","#d1e5f0","#67a9cf", "#67a9cf", "#2166ac"), na.value = "grey98",limits = c(-0, 40))+ 
  theme_classic()+
   theme(axis.text.x=element_text(vjust=0.5, hjust=0.95,size=12),
        plot.title = element_text(margin = margin(t = 10, b = 10), hjust=0.5),
        panel.background= element_rect(fill=NA, color='black'),
        legend.title = element_blank(),
        axis.text.y = element_text(vjust=0.5, size=12), #making the axis text larger 
        axis.title.x = element_blank(),#making the axis title larger 
        axis.title.y = element_blank(),
        strip.text = element_text(size=12))#making the axis title larger 
freq_fig_up

ggsave(filename="../../output/WGCNA/GO_analysis/freq_fig_up.pdf", plot=freq_fig_up, dpi=300, height=12, units="in", limitsize=FALSE)
```

### Frequency >10 down
```{r}
cal_freq_terms_filtered_down <- merged_down_clean %>%
  filter(Number.of.terms>=10)
  #filter(Calcification.direction=="Down")
cal_freq_terms_filtered_down
```

#### Figure
```{r} 
freq_fig_down<-ggplot(cal_freq_terms_filtered_down, aes(y=Number.of.terms,x=reorder(ParentTerm, Number.of.terms)))+
  #facet_wrap(~Calcification.direction, nrow = 1)+
  geom_point(size=5, alpha=1, pch=21,color="black")+
  geom_segment(aes(x=ParentTerm, xend=ParentTerm, y=0, yend=Number.of.terms)) +
  #geom_hline(yintercept = 0, linetype="solid", color = 'black', size=0.5, show.legend = TRUE)+
  coord_flip()+
  scale_y_continuous(expression(GO~term~counts),limits=c(0,27))+
  #scale_color_manual("Direction.of.flat.origin",values= c("up"="#b2182b","no pattern"="grey","down" ="#67a9cf"))+
  #scale_fill_manual("Direction.of.flat.origin",values= c("up"="#b2182b","no pattern"="grey","down" ="#67a9cf"))+
  scale_fill_gradientn(colours=c("white","#d1e5f0","#92c5de","#4393c3","#2166ac"), na.value = "grey98",limits = c(0, 100))+ 
  #scale_fill_gradientn(colours=c("#fddbc7","#f4a582","#d6604d","#b2182b"), na.value = "grey98",limits = c(10, 40))+ 
 #scale_color_gradientn(colours=c("#b2182b","#fddbc7","white","#d1e5f0","#67a9cf", "#67a9cf", "#2166ac"), na.value = "grey98",limits = c(-0, 40))+ 
  theme_classic()+
   theme(axis.text.x=element_text(vjust=0.5, hjust=0.95, size=12),
        plot.title = element_text(margin = margin(t = 10, b = 10), hjust=0.5),
        panel.background= element_rect(fill=NA, color='black'),
        legend.title = element_blank(),
        axis.text.y = element_text(vjust=0.5, size=12), #making the axis text larger 
        axis.title.x = element_blank(),#making the axis title larger 
        axis.title.y = element_blank(),
        strip.text = element_text(size=12))#making the axis title larger 
freq_fig_down

ggsave(filename="../../output/WGCNA/GO_analysis/freq_fig_down.pdf", plot=freq_fig_down, dpi=300, height=12, units="in", limitsize=FALSE)
```
```{r}
compare_figs<-cowplot::plot_grid(freq_fig_up, freq_fig_down, nrow=2, align="v")
compare_figs
```

### Frequency >20 up
```{r}
cal_freq_terms_filtered_up <- merged_up_clean %>%
  filter(Number.of.terms>=20) %>%
  filter(Calcification.direction=="Up")
cal_freq_terms_filtered_up
```

#### Figure
```{r}
#counts$Direction.of.flat.origin<- factor(counts$Direction.of.flat.origin, levels =c("up","no pattern","down"))
#counts$Module<- factor(counts$Module, levels=c("Blue","Brown","Greenyellow","Cyan","Pink","Magenta","Lightcyan","Midnight blue","Purple","Turquiose","Red","Black"))
freq_fig_up<-ggplot(cal_freq_terms_filtered_up, aes(y=Number.of.terms,x=reorder(ParentTerm, Number.of.terms),group=1))+
  #facet_wrap(~Calcification.direction, nrow = 1)+
  geom_point(size=5, alpha=1, pch=21,color="black")+
  geom_segment(aes(x=ParentTerm, xend=ParentTerm, y=0, yend=Number.of.terms)) +
  geom_hline(yintercept = 0, linetype="solid", color = 'black', size=0.5, show.legend = TRUE)+
  coord_flip()+
  scale_y_continuous(expression(GO~term~counts),limits=c(0,60))+
  #scale_color_manual("Direction.of.flat.origin",values= c("up"="#b2182b","no pattern"="grey","down" ="#67a9cf"))+
  #scale_fill_manual("Direction.of.flat.origin",values= c("up"="#b2182b","no pattern"="grey","down" ="#67a9cf"))+
  scale_fill_gradientn(colours=c("white","#fddbc7","#f4a582","#d6604d","#b2182b"), na.value = "grey98",limits = c(0, 100))+ 
 #scale_color_gradientn(colours=c("#b2182b","#fddbc7","white","#d1e5f0","#67a9cf", "#67a9cf", "#2166ac"), na.value = "grey98",limits = c(-0, 40))+ 
  theme_classic()+
   theme(axis.text.x=element_text(vjust=0.5, hjust=0.95,size=12),
        plot.title = element_text(margin = margin(t = 10, b = 10), hjust=0.5),
        panel.background= element_rect(fill=NA, color='black'),
        legend.title = element_blank(),
        axis.text.y = element_text(vjust=0.5, size=12), #making the axis text larger 
        axis.title.x = element_blank(),#making the axis title larger 
        axis.title.y = element_blank(),
        strip.text = element_text(size=12))#making the axis title larger 
freq_fig_up

ggsave(filename="../../output/WGCNA/GO_analysis/freq_fig_up_gr20.pdf", plot=freq_fig_up, dpi=300, height=12, units="in", limitsize=FALSE)
```

### Frequency >20 down
```{r}
cal_freq_terms_filtered_down <- merged_down_clean %>%
  filter(Number.of.terms>=20)
  #filter(Calcification.direction=="Down")
cal_freq_terms_filtered_down
```

#### Figure
```{r} 
freq_fig_down<-ggplot(cal_freq_terms_filtered_down, aes(y=Number.of.terms,x=reorder(ParentTerm, Number.of.terms)))+
  #facet_wrap(~Calcification.direction, nrow = 1)+
  geom_point(size=5, alpha=1, pch=21,color="black")+
  geom_segment(aes(x=ParentTerm, xend=ParentTerm, y=0, yend=Number.of.terms)) +
  #geom_hline(yintercept = 0, linetype="solid", color = 'black', size=0.5, show.legend = TRUE)+
  coord_flip()+
  scale_y_continuous(expression(GO~term~counts),limits=c(0,27))+
  #scale_color_manual("Direction.of.flat.origin",values= c("up"="#b2182b","no pattern"="grey","down" ="#67a9cf"))+
  #scale_fill_manual("Direction.of.flat.origin",values= c("up"="#b2182b","no pattern"="grey","down" ="#67a9cf"))+
  scale_fill_gradientn(colours=c("white","#d1e5f0","#92c5de","#4393c3","#2166ac"), na.value = "grey98",limits = c(0, 100))+ 
  #scale_fill_gradientn(colours=c("#fddbc7","#f4a582","#d6604d","#b2182b"), na.value = "grey98",limits = c(10, 40))+ 
 #scale_color_gradientn(colours=c("#b2182b","#fddbc7","white","#d1e5f0","#67a9cf", "#67a9cf", "#2166ac"), na.value = "grey98",limits = c(-0, 40))+ 
  theme_classic()+
   theme(axis.text.x=element_text(vjust=0.5, hjust=0.95, size=12),
        plot.title = element_text(margin = margin(t = 10, b = 10), hjust=0.5),
        panel.background= element_rect(fill=NA, color='black'),
        legend.title = element_blank(),
        axis.text.y = element_text(vjust=0.5, size=12), #making the axis text larger 
        axis.title.x = element_blank(),#making the axis title larger 
        axis.title.y = element_blank(),
        strip.text = element_text(size=12))#making the axis title larger 
freq_fig_down

ggsave(filename="../../output/WGCNA/GO_analysis/freq_fig_down_gr20.pdf", plot=freq_fig_down, dpi=300, height=12, units="in", limitsize=FALSE)
```

```{r}
compare_figs<-cowplot::plot_grid(freq_fig_up, freq_fig_down, nrow=2, align="v")
compare_figs
```

### All terms up
```{r}
cal_freq_terms_filtered_up_all <- merged_up_clean %>%
  filter(Calcification.direction=="Up")
cal_freq_terms_filtered_up_all
```

#### Figure
```{r}
#counts$Direction.of.flat.origin<- factor(counts$Direction.of.flat.origin, levels =c("up","no pattern","down"))
#counts$Module<- factor(counts$Module, levels=c("Blue","Brown","Greenyellow","Cyan","Pink","Magenta","Lightcyan","Midnight blue","Purple","Turquiose","Red","Black"))
freq_fig_up<-ggplot(cal_freq_terms_filtered_up_all, aes(y=Number.of.terms,x=reorder(ParentTerm, Number.of.terms),group=1))+
  #facet_wrap(~Calcification.direction, nrow = 1)+
  geom_point(size=5, alpha=1, pch=21,color="black")+
  geom_segment(aes(x=ParentTerm, xend=ParentTerm, y=0, yend=Number.of.terms)) +
  geom_hline(yintercept = 0, linetype="solid", color = 'black', size=0.5, show.legend = TRUE)+
  coord_flip()+
  scale_y_continuous(expression(GO~term~counts),limits=c(0,60))+
  #scale_color_manual("Direction.of.flat.origin",values= c("up"="#b2182b","no pattern"="grey","down" ="#67a9cf"))+
  #scale_fill_manual("Direction.of.flat.origin",values= c("up"="#b2182b","no pattern"="grey","down" ="#67a9cf"))+
  scale_fill_gradientn(colours=c("white","#fddbc7","#f4a582","#d6604d","#b2182b"), na.value = "grey98",limits = c(0, 100))+ 
 #scale_color_gradientn(colours=c("#b2182b","#fddbc7","white","#d1e5f0","#67a9cf", "#67a9cf", "#2166ac"), na.value = "grey98",limits = c(-0, 40))+ 
  theme_classic()+
   theme(axis.text.x=element_text(vjust=0.5, hjust=0.95,size=12),
        plot.title = element_text(margin = margin(t = 10, b = 10), hjust=0.5),
        panel.background= element_rect(fill=NA, color='black'),
        legend.title = element_blank(),
        axis.text.y = element_text(vjust=0.5, size=12), #making the axis text larger 
        axis.title.x = element_blank(),#making the axis title larger 
        axis.title.y = element_blank(),
        strip.text = element_text(size=12))#making the axis title larger 
freq_fig_up

ggsave(filename="../../output/WGCNA/GO_analysis/freq_fig_up_ALL.pdf", plot=freq_fig_up, dpi=300, height=17, units="in", limitsize=FALSE)
```

### All terms down
```{r}
cal_freq_terms_filtered_down_all <- merged_down_clean %>%
  filter(Calcification.direction=="Down")
cal_freq_terms_filtered_down_all
```

#### Figure
```{r}
freq_fig_down<-ggplot(cal_freq_terms_filtered_down_all, aes(y=Number.of.terms,x=reorder(ParentTerm, Number.of.terms)))+
  #facet_wrap(~Calcification.direction, nrow = 1)+
  geom_point(size=5, alpha=1, pch=21,color="black")+
  geom_segment(aes(x=ParentTerm, xend=ParentTerm, y=0, yend=Number.of.terms)) +
  #geom_hline(yintercept = 0, linetype="solid", color = 'black', size=0.5, show.legend = TRUE)+
  coord_flip()+
  scale_y_continuous(expression(GO~term~counts),limits=c(0,27))+
  #scale_color_manual("Direction.of.flat.origin",values= c("up"="#b2182b","no pattern"="grey","down" ="#67a9cf"))+
  #scale_fill_manual("Direction.of.flat.origin",values= c("up"="#b2182b","no pattern"="grey","down" ="#67a9cf"))+
  scale_fill_gradientn(colours=c("white","#d1e5f0","#92c5de","#4393c3","#2166ac"), na.value = "grey98",limits = c(0, 100))+ 
  #scale_fill_gradientn(colours=c("#fddbc7","#f4a582","#d6604d","#b2182b"), na.value = "grey98",limits = c(10, 40))+ 
 #scale_color_gradientn(colours=c("#b2182b","#fddbc7","white","#d1e5f0","#67a9cf", "#67a9cf", "#2166ac"), na.value = "grey98",limits = c(-0, 40))+ 
  theme_classic()+
   theme(axis.text.x=element_text(vjust=0.5, hjust=0.95, size=12),
        plot.title = element_text(margin = margin(t = 10, b = 10), hjust=0.5),
        panel.background= element_rect(fill=NA, color='black'),
        legend.title = element_blank(),
        axis.text.y = element_text(vjust=0.5, size=12), #making the axis text larger 
        axis.title.x = element_blank(),#making the axis title larger 
        axis.title.y = element_blank(),
        strip.text = element_text(size=12))#making the axis title larger 
freq_fig_down

ggsave(filename="../../output/WGCNA/GO_analysis/freq_fig_down_ALL.pdf", plot=freq_fig_down, dpi=300, height=17, units="in", limitsize=FALSE)
```

```{r}
compare_figs<-cowplot::plot_grid(freq_fig_up, freq_fig_down, ncol=2, align="h")
compare_figs
```