---
title: "glmmSeq, with pOverA filtering and outliers removed, factor levels reversed"
author: "Zoe Dellaert"
date: "3/29/2023"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is a Rmd file analyzing our raw count data by the glmmSeq package as described in the [vignette](https://cran.r-project.org/web/packages/glmmSeq/vignettes/glmmSeq.html) and [manual](https://cran.r-project.org/web/packages/glmmSeq/glmmSeq.pdf).

```{r}
sessionInfo() #provides list of loaded packages and version of R. I still have version 4.1 for now.
```

First, download the glmmSeq package. I had to install qvalue using BiocManager as described in the [vignette](https://cran.r-project.org/web/packages/glmmSeq/vignettes/glmmSeq.html)

```{r install packages, eval=FALSE}
#if (!requireNamespace("BiocManager", quietly = TRUE))
    #install.packages("BiocManager")
    #BiocManager::install("qvalue"))

#install.packages("glmmSeq") #from CRAN
#devtools::install_github("myles-lewis/glmmSeq") # from github
#install.packages("plotly")
```

Next, load the package 
  
```{r}
library(glmmSeq)
library(dplyr) #load dplyr, which is required to run this tutorial but not included in the vignette
library(kableExtra) #load kableExtra, which is also required to run this tutorial but not included in the vignette
library(plotly) #library plotly to make interactive plots
sessionInfo()
```


and load in raw count data
```{r}
cts_raw <- read.csv("../../../TagSeq_Output/HeronPdam_gene_count_matrix.csv") #load in data
rownames(cts_raw) <- cts_raw[,1] #set first column that contains gene names as rownames
cts_raw <- cts_raw[,-1] #remove the column with gene names
```

Clean up sample names from "RF13B_S85_ALL.bam.gtf" to "RF13B"
```{r}
colnames(cts_raw)<-gsub("_[^_]+$", "",colnames(cts_raw)) #get rid of "_ALL.bam.gtf"
colnames(cts_raw)<-gsub("_[^_]+$", "",colnames(cts_raw)) #get rid of "_S85"
head(colnames(cts_raw)) #see first 6 clean sample names
```

Metadata from this dataset
```{r}
coldata <- read.csv("../../../../RAnalysis/Data/RNA Submission Sample List metadata.csv") #read in metadata file
coldata <- plyr::rename(coldata, c("Sample.Name"="Coral_ID")) #Make a column that represents the colonies
coldata$Colony <- gsub("A", "", coldata$Coral_ID) #ID which sample is from which colony by remobing letters after colony name
coldata$Colony <- gsub("B", "", coldata$Colony) #ID which sample is from which colony by remobing letters after colony name
coldata$Colony <- gsub("C", "", coldata$Colony) #ID which sample is from which colony by remobing letters after colony name
coldata$Colony <- gsub("D", "", coldata$Colony) #ID which sample is from which colony by remobing letters after colony name

coldata$Origin <- factor(coldata$Origin, levels = c("Slope","Flat")) #set variables to factors, with Slope as the baseline
coldata$Colony <- factor(coldata$Colony) #set variables to factors
coldata$Treatment <- factor(coldata$Treatment) #set variables to factors
head(coldata)
```

Data sanity checks:
```{r}
all(rownames(coldata$Coral_ID) %in% colnames(cts_raw)) #are all of the sample names (rows) in the metadata column names in the gene count matrix?
all(rownames(coldata$Coral_ID) == colnames(cts_raw)) #are they the same in the same order?
```

## Remmoving outlier "RF16A" and "RF16C"

```{r}
cts_raw_outrm <- cts_raw %>% select(-c("RF16A","RF16C")) #remove those columns from count matrix
coldata_outrm <- coldata %>% filter(Coral_ID != c("RF16A","RF16C")) #removed those rows from metadata
```

## Using filtered data instead of raw counts as input

```{r}
cts_filt_outrm <- cts_raw_outrm[rowSums(!as.matrix(cts_raw_outrm)) < ncol(cts_raw_outrm), ] #here we remove all genes that were not expressed in any of our samples- we are left with 24,220 genes.
```


## pOverA filtering

```{r}
library(genefilter)

#use pOverA filtering to reduce dataset

ffun<-filterfun(pOverA(0.25,10))  #set up filtering parameters
cts_filt_outrm_poa <- genefilter((cts_filt_outrm), ffun) #apply filter
sum(cts_filt_outrm_poa) #count number of genes left

cts_filt_outrm_poa <- cts_filt_outrm[cts_filt_outrm_poa,] #keep only rows that passed filter
```

After removing outliers, there are now 9012 genes in the filtered dataset.

Data sanity checks:
```{r}
all(rownames(coldata_outrm$Coral_ID) %in% colnames(cts_filt_outrm_poa)) #are all of the sample names (rows) in the metadata column names in the gene count matrix?
all(rownames(coldata_outrm$Coral_ID) == colnames(cts_filt_outrm_poa)) #are they the same in the same order?
```


### Time to calculate dispersion!

"Using negative binomial models requires gene dispersion estimates to be made. This can be achieved in a number of ways. A common way to calculate this for gene i is to use the equation:

Dispersioni = (variancei - meani)/meani2

Starting with raw counts, so using [DESeq2](http://bioconductor.org/packages/devel/bioc/vignettes/DESeq2/inst/doc/DESeq2.html), but could also import a normalized gene count matrix (i.e. transcripts per million) and calculate disperson manually as in the [vignette](https://cran.r-project.org/web/packages/glmmSeq/vignettes/glmmSeq.html).

```{r, eval=FALSE}
library(DESeq2) #library in DEseq to use to calculate dispersions

dds_filt_outrm <- DESeqDataSetFromMatrix(countData = cts_filt_outrm_poa,
                              colData = coldata_outrm,
                              design = ~1) #this is what the glmmseq vignette says, cannot tell if we should put our formula in here yet or just calculate around the intercept.

# dds <- DESeqDataSetFromMatrix(countData = cts_raw,
#                               colData = coldata,
#                               design = ~ Treatment * Origin + (1 | Colony)) #ALTERNATIVE, but probably without the "1 | Colony" bit.

dds_filt_outrm <- DESeq(dds_filt_outrm) #apply DEseq dispersion calculation
dispersions_filt_outrm <- setNames(dispersions(dds_filt_outrm), rownames(cts_filt_outrm_poa)) #save dispersions as vector for each gene

rm(dds_filt_outrm) #remove large DEseq object, we no longer need it once we have saved the dispersions
```


### Size Factors
There is also an option to include size factors for each gene.

```{r, eval=FALSE}
sizeFactors_filt <- estimateSizeFactorsForMatrix(cts_filt_outrm_poa)
```

## Fitting Models

In this case study we want to use time and response as fixed effects and the patients as random effects:

> gene expression ~ Origin * Treatment + (1 | Colony)

To fit this model for all genes we can use the glmmSeq function.

```{r, eval=FALSE}
results <- glmmSeq(~ Treatment * Origin + (1 | Colony),
                   countdata = cts_filt_outrm_poa,
                   metadata = coldata_outrm,
                   dispersion = dispersions_filt_outrm,
                   progress = TRUE)

#this fitting takes a while, so I save it as an RDS and load this in when needed instead of re-running everytime (unless something upstream of here changes)

saveRDS(results, file = "glmmSeq.rds")  #save RDS
```

Showed errors in one gene, Pocillopora_acuta_HIv2___RNAseq.g5950.t1.

```{r}
results <- readRDS(file = "glmmSeq.rds") #load in RDS from previous step / previous iteration

results@errors   # show detailed gene error
```

Interrogate that gene's counts to try to understand this error better.

```{r}
cts_filt_outrm_poa %>% filter(row.names(cts_filt_outrm_poa) %in% c('Pocillopora_acuta_HIv2___RNAseq.g5950.t1')) %>% unlist
```


## Hypothesis testing: How much does the interaction factor affect the fit of the model

I am a little lost here.

```{r, eval=FALSE}
glmmLRT <- glmmSeq(~ Treatment * Origin + (1 | Colony),
                   reduced = ~ Treatment + Origin + (1 | Colony), #changed * to + to have reduced model without interaction term
                   countdata = cts_filt_outrm_poa,
                   metadata = coldata,
                   dispersion = dispersions_filt_outrm, verbose = FALSE)

#this fitting takes a while, so I save it as an RDS and load this in when needed instead of re-running everytime (unless something upstream of here changes)

saveRDS(glmmLRT, file = "glmmLRT.rds") #save RDS
```


```{r, eval=FALSE}
glmmLRT <- readRDS(file = "glmmLRT.rds") #load in RDS from previous step / previous iteration

summary(glmmLRT, "Pocillopora_acuta_HIv2___TS.g25814.t1") #see example statistics for one example gene

summary(results, "Pocillopora_acuta_HIv2___TS.g25814.t1") #see example statistics for one example gene

rm(glmmLRT) #remove this large object, it will not be used again (and you can always reload the saved RDS)
```


### Outputs
```{r}
results@modelData #show model data, basically the different conditions tested
```

### Q values

```{r}
results <- glmmQvals(results) # adds a matrix to the results object containing Q-vals
```


Model fit statistics, w/ p-value ordered by Origin:Treatment:
```{r}
stats <- summary(results)  #save the summary table from the model fit results

kable(stats[order(stats[, 'Treatment:Origin']), ]) %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "400px") #View stats table as a scrollable markdown table

write.csv(stats, "signif_genes.csv") #save stats file as csv for downstream analysis
```


Summary stats for a specific gene. (basically a pretty way of looking at one row of the stats table via "stats["Pocillopora_acuta_HIv2___TS.g25814.t1",]")

```{r}
summary(results, gene = "Pocillopora_acuta_HIv2___TS.g25814.t1")
```

Estimated means based on each gene’s fitted model to show fixed effects and their 95% confidence intervals can be seen in the @predict slot:

```{r}
predict = data.frame(results@predict) #estimated means based on fitted model with LCI (lower) and UCI (upper) confidence intervals

kable(predict) %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "400px")
```


### Q values

```{r}
results <- glmmQvals(results) # adds a matrix to the results object containing Q-vals
```

## Model Plots

As written the plotting function from this package does not work when x1var, which in our case is Treatment, is a factor.

I changed the function to allow for discrete values on the X axis:

```{r}
source(file = "Factor_ggmodelPlot.R")

#I changed their code a lot and need to document all the changes

formPlot(results,geneName = "Pocillopora_acuta_HIv2___RNAseq.g13538.t2",x1var = "Treatment",x2var="Origin",x2shift=NULL) #the plotting code is based of off these matrices, made using their function "formPlot"

plotColours <- c("skyblue","mediumseagreen")
modColours <- c("dodgerblue3","seagreen4")
```

### most significant gene by Treatment:Origin:
```{r}
Factor_ggmodelPlot(results,
            geneName = "Pocillopora_acuta_HIv2___RNAseq.g14176.t1",
            x1var = "Treatment",
            x2var="Origin", addBox = T,
            xlab = "Treatment and Origin",
            colours = plotColours,
            lineColours = plotColours, 
            modelColours = modColours,
            modelSize = 3)
```

To make these plots interactive, you can use ggplotly(_plot code_, tooltip="id")

### most significant gene by Treatment:
```{r}
Factor_ggmodelPlot(results,
            geneName = "Pocillopora_acuta_HIv2___RNAseq.g13538.t2",
            x1var = "Treatment",
            x2var="Origin", addBox = T,
            xlab = "Treatment and Origin",
            colours = plotColours,
            lineColours = plotColours, 
            modelColours = modColours,
            modelSize = 3)
```

### most significant gene by Origin:
```{r}
Factor_ggmodelPlot(results,
            geneName = "Pocillopora_acuta_HIv2___RNAseq.g16163.t1",
            x1var = "Treatment",
            x2var="Origin", addBox = T,
            xlab = "Treatment and Origin",
            colours = plotColours,
            lineColours = plotColours, 
            modelColours = modColours,
            modelSize = 3)
```


```{r}
fcPlot(results, x1var = "Origin", x2var = "Treatment", graphics = "plotly",
       pCutoff = 0.05, useAdjusted = TRUE)
```

```{r}
fcPlot(results, x1var = "Treatment", x2var = "Origin", graphics = "plotly",
       pCutoff = 0.05, useAdjusted = TRUE)
```

## MA Plots

Fixed the script for the maplots , just added "scale_y_continuous(limits = c(-4.2,4.2), breaks = seq(-4,4, 2))" to make the scale the same on both plots.
```{r}
source("fixed_ylim_maPlot.R")
maPlots_filt_outrm <- fixed_ylim_maPlot(results,
                  x1var="Treatment",
                  x2var="Origin",
                  x2Values=c("Flat", "Slope"),
                  graphics="ggplot")

maPlots_filt_outrm$combined
```
